---
title: "ohbm_results"
output: html_document
date: '2023-11-12'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

single script to parse outputs & get results (condensed from `ukb_sim.site_analyses.Rmd`)

load info
```{r}
data_path <- "./ukb_basic"

### LOAD ###
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
```

# Main Analyses

## CENTILES
Do centile scores change based on combat version?

```{r}
pred.csvs <- list.files(path = data_path, pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
  
  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
}
```

```{r, eval=FALSE}
#calculate abs error
cent_abs.diffs <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop raw

saveRDS(cent_abs.diffs, "data/centile_abs_diffs.rds")
```

```{r}
cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
```

Test for differences in abs. error distributions across combat configurations using nonparametric kruskal-wallace test (instead of anova. good that it's nonparametric b/c non-normal distributions BUT also the data is paired!). **confirm with andrew**
Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

**need to confirm w andrew then update output table formatting as needed - skipping for now**

```{r}
#initalize empty dfs to store outputs
sum.df <- data.frame(pheno = character(),
                     term = character(),
                     df = double(),
                     sumsq = double(),
                     meansq = double(),
                     statistic = double(),
                     p.value = double())
t.df <- data.frame("pheno" = character(),
                   "group1" = character(),
                   "group2" = character(),
                   "p.value" = double())
pheno_diff.list <- paste0("abs.diff_", pheno_list)

#run tests
attach(cent_abs.diffs)
for (pheno in pheno_diff.list) {
  # df.aov <- tidy(kruskal.test(cent_abs.diffs[[pheno]] ~ cent_abs.diffs[["dataset"]]))
  # df.aov$pheno <- pheno
  # sum.df <- rbind(sum.df, df.aov)
  
  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=cent_abs.diffs[[pheno]], g=cent_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  t.df <- rbind(t.df, df.pairwise.t)
}
detach()

#apply fdr correction
# sum.df <- sum.df %>%
#   dplyr::filter(term != "Residuals") %>%
#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

t.df <- t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("abs.diff_", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))

# t.df.sum <- t.df %>%
#   group_by(comp) %>%
#   summarize(n_sig_regions = sum(sig_fdr),
#             n_compare = n(),
#             min_pval = min(p.val_fdr),
#             max_pval = max(p.val_fdr))
# t.df.sum
```

Little plot
```{r}
mean.diffs.long <- cent_abs.diffs %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
```

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")
cent.diffs.plot

#ggsave("figs/mean_centile_diffs.jpeg", cent.diffs.plot)
```

## Z-SCORES

look for skew in distribution

```{r}
z_abs.diffs <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "refA")) %>% #drop ref site results
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop raw

#initalize empty dfs to store outputs
# z.sum.df <- data.frame(pheno = character(),
#                        term = character(),
#                        df = double(),
#                        sumsq = double(),
#                        meansq = double(),
#                        statistic = double(),
#                        p.value = double())
z.t.df <- data.frame("pheno" = character(),
                     "group1" = character(),
                     "group2" = character(),
                     "p.value" = double())

z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")

#run tests
attach(z_abs.diffs)
for (pheno in z.pheno_abs.diff.list) {
  # df.aov <- tidy(kruskal.test(z_abs.diffs[[pheno]] ~ z_abs.diffs[["dataset"]]))
  # df.aov$pheno <- pheno
  # z.sum.df <- rbind(z.sum.df, df.aov)
  
  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=z_abs.diffs[[pheno]], g=z_abs.diffs[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  z.t.df <- rbind(z.t.df, df.pairwise.t)
}
detach()

#apply fdr correction
# z.sum.df <- z.sum.df %>%
#   dplyr::filter(term != "Residuals") %>%
#   mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

z.t.df <- z.t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("abs.diff_", "", pheno),
         pheno = sub(".z", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))

# sum_z.t.df <- z.t.df %>%
#   group_by(comp) %>%
#   summarize(n_sig_regions = sum(sig_fdr),
#             n_compare = n(),
#             min.pval = min(p.val_fdr),
#             max.pval = max(p.val_fdr))
# 
# sum_z.t.df
```


## SCALE EFFECTS OF SEX
fix age, get diff between male and female pred. variance
```{r}
pred.csvs <- list.files(path = data_path, pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File),
           sex_standard = sex_effect/f.var)
  
  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset),
         dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)

#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")

plot.info <- sigma.sex.df %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
  unique() %>%
  mutate(dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset))

combined_var <- inner_join(combined_var, plot.info)
```

```{r}
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
var.plot

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
```

```{r}
#plot global
var.plot.global.df <- combined_var %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
var.plot.global
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))

ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```

# Replication 
Permuting site assignments 3x

```{r}
#path to csvs scp'd from cubic
permutation_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_perm_csvs"
```

## CENTILES
Do centile scores change based on combat version?

```{r}
n_perm_list <- lapply(c(1:10), str_pad, 3, pad = "0")
n_perm_list <- paste0("perm-", n_perm_list)
pheno_list.z <- paste0(pheno_list, ".z")

# combined_preds.p <- data.frame()
get.predictions <- function(x){
  df <- data.frame() #new empty dataframe
  pred.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_predictions.csv"), full.names = TRUE)
  for (file in pred.csvs.p) {
    # Read each CSV file
    data <- fread(file)

    # Add a "Source_File" column with the file name
    data <- data %>%
      mutate(Source_File = as.factor(basename(file)),
             perm = as.factor(x)) %>%
      mutate(dataset = gsub("_data|_predictions.csv|perm-|[0-9]|-", "", Source_File))

    # Bind the data to the combined dataframe
    df <- bind_rows(df, data, .id = "File_ID")
  }
  return(df)
}

perm_predictions <- mclapply(n_perm_list, get.predictions, mc.preschedule = FALSE)
```

```{r}
get.abs.diffs <- function(x){
  df <- x %>%
    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
    mutate(dataset = relevel(dataset, ref= "raw")) %>%
    arrange(dataset) %>%
    group_by(participant) %>%
    dplyr::mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #centile diffs
    dplyr::select(-any_of(pheno_list)) %>% #drop
    dplyr::mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>% #z-score diffs
    dplyr::select(-any_of(pheno_list.z)) %>% #drop
    ungroup() %>%
    dplyr::filter(dataset != "raw") #drop raw
  return(df)
}

```

```{r}
perm_diffs <- mclapply(perm_predictions, get.abs.diffs, mc.preschedule = FALSE)
saveRDS(perm_diffs)
```

```{r}
perm_diffs <- readRDS()
```

add ANOVA
```{r}

```

Pairwise T tests

```{r}
#define function
pheno_diff.list <- paste0("abs.diff_", pheno_list)

centile.t.tests <- function(df){
  #initialize empty dfs to store outputs
  t.df <- data.frame("pheno" = character(),
                     "group1" = character(),
                     "group2" = character(),
                     "p.value" = double())
  #run tests
  attach(df)
  for (pheno in pheno_diff.list) {
    df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none"))
    df.pairwise.t$pheno <- pheno
    t.df <- rbind(t.df, df.pairwise.t)
  }
  n_comp <- nrow(t.df)

  #apply fdr correction
  t.df <- t.df %>%
    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                               p.val_fdr >= 0.05 ~ FALSE),
           pheno = sub("abs.diff_", "", pheno)) %>%
    unite(comp, c("group1", "group2")) %>%
   dplyr::mutate(comp = as.factor(comp))

  t.df.sum <- t.df %>%
    group_by(comp) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
  return(t.df.sum)
  #detach()
}
```


```{r}
#apply
perm_t.tests <- mclapply(perm_diffs, centile.t.tests, mc.preschedule = FALSE)
names(perm_t.tests) <- n_perm_list
cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")
```

```{r}
cent_t.tests.p %>%
  group_by(comp) %>%
  dplyr::summarize(n_sig_min=min(n_sig_regions),
            n_sig_max = max(n_sig_regions),
            n_sig_mean = mean(n_sig_regions))
```

Little plots
```{r}
#define plotting functions
plot.cent.diffs <- function(df){
mean.diffs.long <- df %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "raw") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")

print(cent.diffs.plot)
}
```

```{r}
lapply(perm_diffs, plot.cent.diffs)
```

W/in subj average plots
```{r}
mean.subj.centile.err <- function(df){
mean.diffs.subj <- df %>%
  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  rowwise() %>%
  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))

return(mean.diffs.subj)
#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
}
perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
names(perm_cent.err) <- n_perm_list
cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")

#write out for plotting later
saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
```

```{r}
cent.diffs.v.plot <- cent_err.p %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```


## Z-SCORES

*need to look at distribution to check for skew!*

```{r}
#within each phenotype, iterate paired t test between each combo of combat dfs 
pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")

z.t.tests <- function(df){
  #initialize empty dfs to store outputs
  t.df <- data.frame("pheno" = character(),
                     "group1" = character(),
                     "group2" = character(),
                     "p.value" = double())
  #run tests
  attach(df)
  for (pheno in pheno_diff.list.z) {
    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
    df.pairwise.t$pheno <- pheno
    t.df <- rbind(t.df, df.pairwise.t)
  }
  n_comp <- nrow(t.df)

  #apply fdr correction
  t.df <- t.df %>%
    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                               p.val_fdr >= 0.05 ~ FALSE),
           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
    unite(comp, c("group1", "group2")) %>%
   dplyr::mutate(comp = as.factor(comp))

  t.df.sum <- t.df %>%
    group_by(comp) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
  return(t.df.sum)
  #detach()
}

perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
names(perm_t.tests_z) <- n_perm_list
z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
```

```{r}
fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
```

```{r}
z_t.tests.p %>%
  group_by(comp) %>%
  dplyr::summarize(n_sig_min=min(n_sig_regions),
            n_sig_max = max(n_sig_regions),
            n_sig_mean = mean(n_sig_regions))
```

## SCALE EFFECTS OF SEX
fix age, get diff between male and female pred. variance
```{r}
# combined_preds.p <- data.frame()
get.var <- function(x){
  df <- data.frame() #new empty dataframe
  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
  for (file in var.csvs.p) {
    # Read each CSV file
    data <- fread(file)

    # Add a "Source_File" column with the file name
    data <- data %>%
      mutate(Source_File = as.factor(basename(file)),
             perm = as.factor(x)) %>%
      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
             sex_standard = sex_effect/f.var,
             pheno = as.character(pheno))

    # Bind the data to the combined dataframe
    df <- bind_rows(df, data, .id = "File_ID")
  }
  return(df)
}

perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
perm_var.all <- bind_rows(perm_var, .id = "column_label")
```

maybe better to get sex_summary.csv - just need signifigance

```{r}
#get summary df for p vals, plotting
sigma.sex.df <- fread("derivatives/ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
         dataset = gsub("-data", "", dataset),
         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
         dataset = gsub("-", "", dataset),
         #dataset = as.character(dataset),
         pheno = as.character(pheno)) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)

perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
```

```{r}
  sex_effect_plot <- perm_var_info %>%
    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
    ggplot() +
    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
    facet_wrap(~perm, ncol=5) +
    ggtitle(paste("Standardized effect of male sex on variance")) +
    theme(axis.text.x = element_blank())
  
    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
    print(sex_effect_plot)
```


# Secondary: Vary M:F

```{r}
ratio_data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"
raw_files <- list.files(path = ratio_data_path, pattern = "_pred.csv", full.names = TRUE)

df_list <- list() #new empty list
for (file in raw_files) {
  # Read each CSV file
    data <- fread(file)

    # Add a "Source_File" column with the file name
    data <- data %>%
      mutate(Source_File = as.factor(basename(file))) %>%
      mutate(dataset = gsub("_data|_raw.csv|prop-|[0-9]|-", "", Source_File))

    # Bind the data to the combined dataframe
    df_list <- c(df_list, list(data))
    assign("ratio_sites_list", df_list)
}
length(ratio_sites_list)
```


## CENTILES
Do centile scores change based on combat version?

need to look and see what happens for subjects who aren't in all samples!

Little plots
```{r}
lapply(ratio_sites_list, plot.cent.diffs)
```

Plot boxplots to compare across prop.
```{r}
ratio_diffs.df <- bind_rows(ratio_diffs) %>% 
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(perm, dataset) %>%
  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

ratio_diffs.df %>%
  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
  ggplot() +
    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
```

get w/in subj mean magnitude centile error

```{r}
ratio_subj_diffs.df <- bind_rows(ratio_diffs) %>% 
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(perm, dataset) %>%
  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

ratio_diffs.df %>%
  mutate(centile_abs.diff = centile_abs.diff * 100) %>%
  ggplot() +
    geom_boxplot(aes(x=perm, y=centile_abs.diff, fill=dataset)) +
  labs(y= "Magnitude Centile Error", x = "M:F Ratio in Site B") +
  ggtitle("Within-feature mean magnitude of error in centile scores across ComBat \nconfigurations as sex imbalance changes")
```

W/in subj average plots
```{r, eval=FALSE}
mean.subj.centile.err <- function(df){
mean.diffs.subj <- df %>%
  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  rowwise() %>%
  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))

return(mean.diffs.subj)
#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
}
ratio_diffs.err <- mclapply(ratio_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
names(ratio_diffs.err) <- levels(ratio_diffs.df$perm)
subj.cent_err.r <- bind_rows(ratio_diffs.err, .id = "proportion")

#write out for plotting later
saveRDS(subj.cent_err.r, "derivatives/ukb_ratio/subj_centile_errors.RDS")
```

```{r}
subj.cent_err.r <- readRDS("derivatives/ukb_ratio/subj_centile_errors.RDS")
```

```{r}
cent.diffs.v.plot.r <- subj.cent_err.r %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=proportion, y=mean, fill=dataset)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
ggsave(filename="figs/ratio_centile_violin.png", cent.diffs.v.plot.r)
```

```{r}
cent.diffs.v.boxplot.r <- subj.cent_err.r %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=proportion, y=mean, fill=dataset)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
ggsave(filename="figs/ratio_centile_box.png", cent.diffs.v.plot.r)
```

### Sex-Biases

```{r}
cent.diffs.v.boxplot.r_sex_free <- subj.cent_err.r %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
  facet_wrap(~dataset, scale="free_y") +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
ggsave(filename="figs/ratio_centile_box_sex_free.png", cent.diffs.v.boxplot.r_sex_free)
```


```{r}
#trying to add significance testing
cent.diffs.by.sex.stat.test <- subj.cent_err.r %>%
  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
  group_by(dataset, proportion) %>%
  wilcox_test(mean ~ sex) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance() %>%
  add_xy_position(x = "proportion")

cent.diffs.v.boxplot.r_sex <- subj.cent_err.r %>%
  mutate(mean = mean*100) %>% #for visuals
  dplyr::filter(dataset != "cf") %>% #because it's the same as cf.lm
  droplevels() %>%
  ggplot() +
  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
  facet_wrap(~dataset) +
  labs(y= "Mean Magnitude Centile Error", x = "Site B M:F") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  stat_pvalue_manual(
    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0, bracket.shorten=0.005,
    #hide.ns = TRUE
    bracket.nudge.y = -0.5
    ) +
  ggtitle("Within-subject error in centile scores across ComBat configurations while varying M:F")
cent.diffs.v.boxplot.r_sex
ggsave(filename="figs/ratio_abs.centile_box_sex.png", cent.diffs.v.boxplot.r_sex, width = 25, height = 12, units = "cm")
```

get diffs (not abs)

```{r}
get.diffs <- function(x){
  df <- x %>%
    mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
    mutate(dataset = relevel(dataset, ref= "raw")) %>%
    arrange(dataset) %>%
    group_by(participant) %>%
    dplyr::mutate(across(all_of(pheno_list), ~ (. - first(.)), .names = "diff_{.col}")) %>% #centile diffs
    dplyr::select(-any_of(pheno_list)) %>% #drop
    dplyr::mutate(across(ends_with(".z"), ~ (. - first(.)), .names = "diff_{.col}")) %>% #z-score diffs
    dplyr::select(-any_of(pheno_list.z)) %>% #drop
    ungroup() %>%
    dplyr::filter(dataset != "raw") #drop raw
  return(df)
}
```

```{r, eval=FALSE}
#get diff in centiles and z scores across ratios
ratio_diffs2 <- mclapply(ratio_predictions, get.diffs, mc.preschedule = FALSE)
length(ratio_diffs2)
```

take w/in subj mean
```{r, eval=FALSE}
mean.subj.centile.err2 <- function(df){
mean.diffs.subj <- df %>%
  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  rowwise() %>%
  dplyr::mutate(mean = mean(c_across(starts_with("diff"))))

return(mean.diffs.subj)
#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
}
ratio_diffs.err2 <- mclapply(ratio_diffs2, mean.subj.centile.err2, mc.preschedule = FALSE)
names(ratio_diffs.err2) <- levels(ratio_diffs.df$perm)
subj.cent_err.r2 <- bind_rows(ratio_diffs.err2, .id = "proportion")
saveRDS(subj.cent_err.r2, file="derivatives/ukb_ratio/ratio_centile_diffs.rds")
```

```{r}
#load
subj.cent_err.r2 <- readRDS("derivatives/ukb_ratio/ratio_centile_diffs.rds")
```

```{r}
cent.diffs.v.boxplot.r_sex2 <- subj.cent_err.r2 %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
  facet_wrap(~dataset) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations while varying M:F")
ggsave(filename="figs/ratio_centile_box_sex2.png", cent.diffs.v.boxplot.r_sex)
```

Test for sex-effects in subj's mean centile errors

```{r}
cent.diffs.by.sex.stat.test <- subj.cent_err.r2 %>%
  group_by(dataset, proportion) %>%
  wilcox_test(mean ~ sex) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance() %>%
  add_xy_position(x = "proportion", dodge = 0.8)

cent.diffs.v.boxplot.r_sex_free2 <- subj.cent_err.r2 %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=proportion, y=mean, fill=sex)) +
  facet_wrap(~dataset, scale="free_y") +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  stat_pvalue_manual(
    cent.diffs.by.sex.stat.test, label = "p.adj.signif", tip.length = 0.01,
    #hide.ns = TRUE
    ) +  
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations while varying M:F")
cent.diffs.v.boxplot.r_sex_free2
ggsave(filename="figs/ratio_centile_box_sex_freescale2.png", cent.diffs.v.boxplot.r_sex_free2, width = 15, height=7)
```

### Extremeness-Biases
Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)

test w/in subj avg centile ~ w/in subj avg centile error

```{r}
#mean w/in subj error
subj.cent_err.only <- subj.cent_err.r2 %>%
  dplyr::select(!starts_with("diff")) %>%
  rename(mean_diff = mean)
```

```{r, eval=FALSE}
#calc mean centile
subj.cent.mean <- function(df){
mean.subj <- df %>%
  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  rowwise() %>%
  dplyr::mutate(mean_cent = mean(c_across(pheno_list))) %>%
  dplyr::select(!pheno_list)

return(mean.subj)
#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
}

ratio_cent_means <- mclapply(ratio_predictions, subj.cent.mean, mc.preschedule = FALSE, mc.cores=12L)
#bind predictions
names(ratio_cent_means) <- n_ratio_list
ratio_cent_means.df <- bind_rows(ratio_cent_means, .id = "column_label")
saveRDS(ratio_cent_means.df, file="derivatives/ukb_ratios/ratio_cent_means.csv")
```

## Z-SCORES

```{r}
#within each phenotype, iterate paired t test between each combo of combat dfs 
pheno_diff.list.z <- paste0("abs.diff_", pheno_list, ".z")

z.t.tests <- function(df){
  #initialize empty dfs to store outputs
  t.df <- data.frame("pheno" = character(),
                     "group1" = character(),
                     "group2" = character(),
                     "p.value" = double())
  #run tests
  attach(df)
  for (pheno in pheno_diff.list.z) {
    tryCatch(df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=df[[pheno]], g=df[["dataset"]], paired = TRUE, p.adj = "none")), error=function(e) { traceback() }) #
    df.pairwise.t$pheno <- pheno
    t.df <- rbind(t.df, df.pairwise.t)
  }
  n_comp <- nrow(t.df)

  #apply fdr correction
  t.df <- t.df %>%
    dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = n_comp),
           sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                               p.val_fdr >= 0.05 ~ FALSE),
           pheno = gsub("abs.diff_|\\.z", "", pheno)) %>%
    unite(comp, c("group1", "group2")) %>%
   dplyr::mutate(comp = as.factor(comp))

  t.df.sum <- t.df %>%
    group_by(comp) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
  return(t.df.sum)
  #detach()
}

perm_t.tests_z <- mclapply(perm_diffs, z.t.tests, mc.preschedule = FALSE)
names(perm_t.tests_z) <- n_perm_list
z_t.tests.p <- bind_rows(perm_t.tests_z, .id = "column_label")
```

```{r}
fwrite(z_t.tests.p, file="ukb_perm_csvs/z_t.tests.csv")
```

```{r}
z_t.tests.p %>%
  group_by(comp) %>%
  dplyr::summarize(n_sig_min=min(n_sig_regions),
            n_sig_max = max(n_sig_regions),
            n_sig_mean = mean(n_sig_regions))
```

## SCALE EFFECTS OF SEX
fix age, get diff between male and female pred. variance
```{r}
# combined_preds.p <- data.frame()
get.var <- function(x){
  df <- data.frame() #new empty dataframe
  var.csvs.p <- list.files(path = permutation_path, pattern = paste0(x, ".+_variance.csv"), full.names = TRUE)
  for (file in var.csvs.p) {
    # Read each CSV file
    data <- fread(file)

    # Add a "Source_File" column with the file name
    data <- data %>%
      mutate(Source_File = as.factor(basename(file)),
             perm = as.factor(x)) %>%
      mutate(dataset = gsub("_data|_variance.csv|perm-|[0-9]|-", "", Source_File),
             sex_standard = sex_effect/f.var,
             pheno = as.character(pheno))

    # Bind the data to the combined dataframe
    df <- bind_rows(df, data, .id = "File_ID")
  }
  return(df)
}

perm_var <- mclapply(n_perm_list, get.var, mc.preschedule = FALSE)
perm_var.all <- bind_rows(perm_var, .id = "column_label")
```

maybe better to get sex_summary.csv - just need signifigance

```{r}
#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_perm_csvs/sex_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(perm = gsub("ukb-CN-data-agefilt-|-raw|-cf|.lm|.gam|.gamlss|-data", "", dataset),
         dataset = gsub("-data", "", dataset),
         dataset = gsub("simsites|ukb-CN-agefilt-|perm-|[0-9]", "", dataset),
         dataset = gsub("-", "", dataset),
         #dataset = as.character(dataset),
         pheno = as.character(pheno)) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE),
         perm = factor(perm, levels = unique(perm), ordered=FALSE)) %>%
  #mutate(dataset = relevel(dataset, ref= "raw")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr, perm, sig.drop_bf.corr)

perm_var_info <- inner_join(perm_var.all, sigma.sex.df)
```

```{r}
  sex_effect_plot <- perm_var_info %>%
    mutate(sex_standard = case_when(sig.drop_bf.corr == TRUE ~ sex_standard,
                                    sig.drop_bf.corr == FALSE ~ 0)) %>% #only significant effects
    ggplot() +
    geom_col(aes(x=pheno, y=sex_standard, fill=dataset)) +
    facet_wrap(~perm, ncol=5) +
    ggtitle(paste("Standardized effect of male sex on variance")) +
    theme(axis.text.x = element_blank())
  
    ggsave(paste0("figs/male_variance_perms.jpeg"), sex_effect_plot)
    print(sex_effect_plot)
```

# Secondary: LBCC Data
