---
title: "hbm_paper_results"
output: html_document
date: '2023-11-12'
---

Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
source("./cubic_scripts/R_scripts/ranked_welch_tests.R")

#ATLAS
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
#write out new dk atlas w/o medial wall
dk_nomed <- dk
dk_nomed$data <- dk_nomed$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")
#plotting info from dk atlas
dk.parc <- dk_nomed$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

#PHENO LISTS
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
pheno_list.z <- paste0(pheno_list, ".z")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

#PATH
der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/lifespan"
```

#Load

Combat-GAM Centiles - study-level batch correction w/o infants

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))
gam_centile_df <- gam_centile_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
gamlss_centile_df <- gamlss_centile_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(gamlss_centile_df)
```

```{r}
min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years)
min(gamlss_centile_df$age_years)
```

Unharmonized Centiles
```{r}
raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
raw_centile_df <- raw_centile_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(raw_centile_df)
min(raw_centile_df$age_years) == min(gamlss_centile_df$age_years)
```

Deltas

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
deltas_df <- deltas_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

dim(deltas_df)
min(deltas_df$age_years) == min(gamlss_centile_df$age_years)
```

Nature paper models to look at age effects in variance
from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)

```{r}
paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
```

#Clean & EDA

Sample plot
```{r}
#reorder studies by age
age.df <- gamlss_centile_df %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_years)) %>%
  arrange(-mean_age)

sample_boxplot <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=study, x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
  labs(color="Sex", y = "Study")

#overall sample
full_boxplot <- 
  gamlss_centile_df %>%
  mutate(ylab = "  Full \n  Sample") %>%
  ggplot(aes(y=ylab, x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x = "Age (Years)", y="")

sample_barplot <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(color="Sex")

full_barplot <- 
  gamlss_centile_df %>%
  mutate(ylab = "Full Sample") %>%
  ggplot() +
  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
  labs(color="Sex", x = "Sex")

ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          align = "hv", 
          common.legend = TRUE,
          widths=c(2, 1),
          heights=c(10, 1))
```

#Batch Effects

Test to see if centile score is associated with study in either ComBatLS, ComBat-GAM, or non-harmonized data.

## Unharmonized
```{r}
#clean predictions
raw_centile_df.clean <- raw_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

skim(raw_centile_df.clean)
```

w/in feature ANOVAs

```{r}
raw.feat_to_test <- intersect(pheno_list, names(raw_centile_df.clean))
length(raw.feat_to_test)

df_raw_feat_aov <- data.frame()
for(pheno in raw.feat_to_test){
  # print(pheno)
  df <- tidy(aov(raw_centile_df.clean[[pheno]] ~ raw_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_raw_feat_aov <- rbind(df_raw_feat_aov, df)
}

df_raw_feat_aov.study <- df_raw_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_raw_feat_aov.study %>%
  ggplot() +
  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)

df_raw_feat_aov.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```
```{r}
raw_centile_df.mean <- raw_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$participant)
```

```{r}
raw_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

raw_centile_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
  theme(legend.position = "none") +
  facet_wrap(~study)
```
linear model

```{r}
summary(lm(mean_cent ~ study, data=raw_centile_df.mean))
summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
raw_z_df.clean <- raw_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
raw.feat_to_test.z <- paste0(raw.feat_to_test, ".z")
df_raw_feat_aov.z <- data.frame()
for(pheno in raw.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(raw_z_df.clean[[pheno]] ~ raw_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_raw_feat_aov.z <- rbind(df_raw_feat_aov.z, df)
}

df_raw_feat_aov.z.study <- df_raw_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_raw_feat_aov.z.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

average w/in subject

```{r}
raw_z_df.mean <- raw_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$participant)
```

```{r}
raw_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
# +
#   facet_wrap(~study)
```

linear model

```{r}
summary(lm(mean_z ~ study, data=raw_z_df.mean))
summary(aov(mean_z ~ study, data=raw_z_df.mean))
```

## ComBat-GAM

```{r}
#clean predictions
gam_centile_df.clean <- gam_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

skim(gam_centile_df.clean)
```

w/in feature ANOVAs

```{r}
gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
length(gam.feat_to_test)

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

df_gam_feat_aov.study <- df_gam_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gam_feat_aov.study %>%
  ggplot() +
  geom_histogram(aes(x=etasq, fill=sig_fdr), alpha=0.5)

df_gam_feat_aov.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

average w/in subject

```{r}
gam_centile_df.mean <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$participant)
```

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

gam_centile_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
  theme(legend.position = "none") +
  facet_wrap(~study)
```

linear model

```{r}
summary(lm(mean_cent ~ study, data=gam_centile_df.mean))
summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
gam_z_df.clean <- gam_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gam_feat_aov.z.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

average w/in subject

```{r}
gam_z_df.mean <- gam_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
```

```{r}
gam_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
# +
#   facet_wrap(~study)
```

linear model

```{r}
summary(lm(mean_z ~ study, data=gam_z_df.mean))
summary(aov(mean_z ~ study, data=gam_z_df.mean))
```


##ComBatLS

```{r}
#clean predictions
gamlss_centile_df.clean <- gamlss_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
```


```{r}
ggplot(gamlss_centile_df.clean) +
  geom_point(aes(x=age_days, y=rh_Vol_insula, color=study), alpha=0.5) +
  theme(legend.position="none")
```

w/in feature ANOVAs

```{r}
gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))

df_gamlss_feat_aov <- data.frame()
for(pheno in gamlss.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gamlss_feat_aov.study %>%
  ggplot() +
  geom_histogram(aes(x=etasq, fill=sig_fdr))

df_gamlss_feat_aov.study %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  nrow()
```



average w/in subject

```{r}
gamlss_centile_df.mean <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            log_age=unique(log_age),
            sexMale=unique(sexMale),
            fs_version=unique(fs_version),
            mean_cent=mean(centile))
nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$participant)
```

```{r}
ggplot(gamlss_centile_df.mean) +
  geom_point(aes(x=(age_days/365.25), y=mean_cent, color=study), alpha=0.5) +
  theme(legend.position="none")

ggplot(gamlss_centile_df.clean) +
  geom_point(aes(x=age_days, y=lh_Vol_inferiortemporal, color=study), alpha=0.2) +
  theme(legend.position="none")
```

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

gamlss_centile_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_cent, color=study), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_cent, color=study), method="lm") +
  theme(legend.position = "none")

gamlss_centile_df.mean %>%
  dplyr::filter(age_days>365) %>%
  ggplot(aes(x=age_days/365.25, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

linear model

```{r}
summary(lm(mean_cent ~ study, data=gamlss_centile_df.mean))
summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
gamlss_z_df.clean <- gamlss_centile_df %>%
  mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
gamlss.feat_to_test.z <- paste0(gamlss.feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gamlss_feat_aov.z.study %>%
  dplyr::filter(sig_fdr == TRUE) %>%
  nrow()
```

average w/in subject

```{r}
gamlss_z_df.mean <- gamlss_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site=unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
```

```{r}
gamlss_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
  facet_wrap(~study) +
  theme(legend.position="none")
```

linear model

```{r}
summary(lm(mean_z ~ study, data=gamlss_z_df.mean))
summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
```

##Comparison Plots

compare magnitude of effect size for study (w/in feature anovas) across unharmonized, combat-gam and combatLS centiles

```{r}
df_gamlss_feat_aov.study$combat <- "gamlss"
df_gam_feat_aov.study$combat <- "gam"
df_raw_feat_aov.study$combat <- "raw"

#put all data frames into list
df_list.tmp <- list(df_gamlss_feat_aov.study, df_gam_feat_aov.study, df_raw_feat_aov.study)

#merge all data frames in list
feat_aov.df <- df_list.tmp %>% reduce(full_join)

feat_aov.df <- feat_aov.df %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

feat_aov.df %>%
  ggplot() +
  geom_point(aes(x=etasq, y=pheno, color=combat, shape=sig_fdr), alpha=0.6) +
  facet_wrap(~pheno_cat, scales="free")
```
```{r}
gamlss_centile_df.mean$combat <- "gamlss"
gam_centile_df.mean$combat <- "gam"
raw_centile_df.mean$combat <- "raw"

#put all data frames into list
df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)

#merge all data frames in list
mean_cents.df <- df_list.tmp %>% reduce(full_join)

mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5)) +
  facet_wrap(~combat)
```

#Feature Centile ~ Mean Study Age

Get study info
```{r}
study_info <- raw_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(study) %>%
  summarize(study_mean_age=mean(age_years),
            study_median_age=median(age_years),
            study_n= n()) %>%
  dplyr::select(study, study_mean_age, study_median_age, study_n)
```

W/in feat lms

##Unharmonized

```{r}
raw_centile_df.clean_study <- full_join(raw_centile_df.clean, study_info)

df_raw_feat_lm <- data.frame()
for(pheno in raw.feat_to_test){
  # print(pheno)
  df <- tidy(aov(raw_centile_df.clean_study[[pheno]] ~ raw_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_raw_feat_lm <- rbind(df_raw_feat_lm, df)
}

df_raw_feat_lm <- df_raw_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_lm %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_raw_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

##ComBat-GAM

```{r}
gam_centile_df.clean_study <- full_join(gam_centile_df.clean, study_info)

df_gam_feat_lm <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean_study[[pheno]] ~ gam_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_gam_feat_lm <- rbind(df_gam_feat_lm, df)
}

df_gam_feat_lm <- df_gam_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

##ComBatLS

```{r}
gamlss_centile_df.clean_study <- full_join(gamlss_centile_df.clean, study_info)

df_gamlss_feat_lm <- data.frame()
for(pheno in gamlss.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean_study[[pheno]] ~ gamlss_centile_df.clean_study[["study_mean_age"]]))
  df$pheno <- pheno
  df_gamlss_feat_lm <- rbind(df_gamlss_feat_lm, df)
}

df_gamlss_feat_lm <- df_gamlss_feat_lm %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_lm %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

df_gamlss_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic))
```

*statistic ~ effect size?

```{r}
df_raw_feat_lm$combat <- "raw"
df_gam_feat_lm$combat <- "gam"
df_gamlss_feat_lm$combat <- "gamlss"

#put all data frames into list
df_list.tmp <- list(df_raw_feat_lm, df_gam_feat_lm, df_gamlss_feat_lm)

#merge all data frames in list
all_feat_lm <- df_list.tmp %>% reduce(full_join)

all_feat_lm %>%
  ggplot() +
  geom_histogram(aes(x=statistic, fill=combat, color=combat), position="identity", alpha=0.5, bins=50)
```

#Subject-Level Mean Centiles ~ Mean Study Age

```{r}
raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))

raw_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```


```{r}
gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))

gam_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
gam_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))

gamlss_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
gamlss_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

Plots

```{r}
gamlss_centile_df.mean2$combat <- "gamlss"
gam_centile_df.mean2$combat <- "gam"
raw_centile_df.mean2$combat <- "raw"

#put all data frames into list
df_list.tmp <- list(gamlss_centile_df.mean2, gam_centile_df.mean2, raw_centile_df.mean2)

#merge all data frames in list
all_subj_means.df <- df_list.tmp %>% reduce(full_join)

all_subj_means.df %>%
  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")

all_subj_means.df %>%
  ggplot(aes(x=study_mean_age, y=mean_cent, color=combat)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm") +
  facet_wrap(~combat)

summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))
```

```{r}
summary(lm(mean_cent ~ study_n, data=raw_centile_df.mean2))
summary(lm(mean_cent ~ study_n, data=gam_centile_df.mean2))
summary(lm(mean_cent ~ study_n, data=gamlss_centile_df.mean2))

all_subj_means.df %>%
  ggplot(aes(x=study_n, y=mean_cent, color=combat)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="lm")
```

#ComBat-GAM vs ComBatLS: Delta Centile

clean delta df

```{r}
cent_deltas_df <- deltas_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

#get subject-level means
mean_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_cent=(mean(abs(centile)))*100) 

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
```

check mean subj delta ~ study mean age

```{r}
delta_lm <- lm(mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(delta_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(delta_lm)[["study_mean_age"]], 
              intercept = coef(delta_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

mean abs delta ~ study mean age

```{r}
abs_lm <- lm(abs_mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(abs_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_lm)[["study_mean_age"]], 
              intercept = coef(abs_lm)[["(Intercept)"]])+
  theme(legend.position="none")

mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

abs_lm2 <- lm(abs_mean_delta_cent ~ study_mean_age + study_n + abs(age_diff), data=mean_cent_deltas_df_study)
summary(abs_lm2)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=abs(age_diff), y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05) +
  geom_smooth(method="lm")
```

*options: rerun from the beginning w/o UKB? Or just control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting?

###Global Features
Look at global features

```{r}
glob_cent_deltas_df <- cent_deltas_df %>%
  dplyr::select(study, age_years, sex, participant, fs_version, paste0("delta.", vol_list_global)) %>%
  mutate(across(paste0("delta.", vol_list_global), abs, .names = "abs_{.col}"))

glob_cent_deltas_df_long <- glob_cent_deltas_df %>%
  pivot_longer(cols=starts_with("abs"), names_to="pheno", values_to = "abs_delta") %>%
  mutate(pheno=gsub("abs_delta.", "", pheno))

glob_cent_deltas_df_long %>%
  ggplot() +
  geom_boxplot(aes(y=abs_delta, color=pheno), outlier.shape = NA) + 
scale_y_continuous(limits = quantile(glob_cent_deltas_df_long$abs_delta, c(0.1, 0.9)))
```
```{r}
glob_cent_deltas_df_study <- full_join(glob_cent_deltas_df, study_info)

abs_gmv_lm <- lm(abs_delta.GMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_gmv_lm)

glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.GMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_gmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_gmv_lm)[["(Intercept)"]])+
  theme(legend.position="none") +
  scale_y_continuous(limits = c(0, 0.1))


abs_sgmv_lm <- lm(abs_delta.sGMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_sgmv_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.sGMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_sgmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_sgmv_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

abs_wmv_lm <- lm(abs_delta.WMV ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)

summary(gam(abs_delta.WMV ~ pb(study_mean_age) + study_n, data=glob_cent_deltas_df_study))

summary(abs_wmv_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.WMV)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_wmv_lm)[["study_mean_age"]], 
              intercept = coef(abs_wmv_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

abs_ven_lm <- lm(abs_delta.Ventricles ~ study_mean_age + study_n, data=glob_cent_deltas_df_study)
summary(abs_ven_lm)
glob_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_delta.Ventricles)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_ven_lm)[["study_mean_age"]], 
              intercept = coef(abs_ven_lm)[["(Intercept)"]])+
  theme(legend.position="none")+
  scale_y_continuous(limits = c(0, 0.1))

glob_cent_deltas_df_long <- full_join(glob_cent_deltas_df_long, study_info)

glob_cent_deltas_df_long <- glob_cent_deltas_df_long %>%
  mutate(age_diff=age_years-study_mean_age)

glob_cent_deltas_df_long %>%
  ggplot(aes(x=study_mean_age, y=abs_delta, color=pheno)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm") +
  facet_wrap(~pheno, scales="free_y")
```

####Compare w/ Nature Paper
Compare to Nature paper age effects in sigma

```{r}
#gmv
paper_fit_gmv$param$sigma$fixef
gmv_b1 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))1"])
gmv_b2 <- unname(paper_fit_gmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 3))2"])

#formula
gmv_age_sig <- parse(text = paste0(gmv_b1, "*x^(-2) + ", gmv_b2,"*x^(3)"))
gmv_age_sig_der <- D(gmv_age_sig,"x")

#sgmv
paper_fit_sgmv$param$sigma$fixef
sgmv_b1 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))1"])
sgmv_b2 <- unname(paper_fit_sgmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2))2"])

sgmv_age_sig <- parse(text = paste0(sgmv_b1, "*x^(-2) + ", sgmv_b2,"*x^(-2)*log(x)"))
sgmv_age_sig_der <- D(sgmv_age_sig,"x")

#wmv
paper_fit_wmv$param$sigma$fixef
wmv_b1 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))1"])
wmv_b2 <- unname(paper_fit_wmv$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, 2))2"])

wmv_age_sig <- parse(text = paste0(wmv_b1, "*x^(-2) + ", wmv_b2,"*x^(s)"))
wmv_age_sig_der <- D(wmv_age_sig,"x")

#ven
paper_fit_ven$param$sigma$fixef
ven_b1 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))1"])
ven_b2 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))2"])
ven_b3 <- unname(paper_fit_ven$param$sigma$fixef["bfpNA(AgeTransformed, powers = c(-2, -2, -2))3"])

ven_age_sig <- parse(text = paste0(ven_b1, "*x^(-2) + ", ven_b2,"*x^(-2)*log(x) + ", ven_b3, "*x^(-2)*log(x)^(2)"))
ven_age_sig_der <- D(ven_age_sig,"x")
```
try summing abs(betas) within each pheno to get overall magnitude of age effects

```{r}
paper_sigma_age <- data.frame("pheno" = c("GMV", "sGMV", "WMV", "Ventricles"),
                              "paper_abs_betas_sum" = c(sum(abs(gmv_b1), abs(gmv_b2)), 
                                                  sum(abs(sgmv_b1), abs(sgmv_b2)),
                                                  sum(abs(wmv_b1), abs(wmv_b2)),
                                                  sum(abs(ven_b1), abs(ven_b2), abs(ven_b3))),
                              "paper_betas_sum" = c(sum(gmv_b1, gmv_b2), 
                                                  sum(sgmv_b1, sgmv_b2),
                                                  sum(wmv_b1, wmv_b2),
                                                  sum(ven_b1, ven_b2, ven_b3)),
                              "lm_beta" = c(coef(abs_gmv_lm)[["study_mean_age"]],
                                            coef(abs_sgmv_lm)[["study_mean_age"]],
                                            coef(abs_wmv_lm)[["study_mean_age"]],
                                            coef(abs_ven_lm)[["study_mean_age"]]),
                              "quick_var_range" = c((38.85552-16.93754), (0.2380905-0.1447285), (29.34474-13.4674), (17.63019-0.3520969))
                              )
paper_sigma_age %>%
  ggplot() +
  geom_col(aes(x=pheno, y=paper_abs_betas_sum, fill=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_col(aes(x=pheno, y=paper_betas_sum, fill=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_point(aes(x=abs(lm_beta), y=paper_betas_sum, color=pheno))

paper_sigma_age %>%
  ggplot() +
  geom_point(aes(x=abs(lm_beta), y=quick_var_range, color=pheno))
```

###W/In Cortical Features

Looking at averages w/in each cortical feature since they have different trajectories of variance across the lifespan (see Nature paper)

####Regional Volumes

```{r}
#get subject-level means
mean_vol_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", vol_list_regions)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_vol_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_vol_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_vol_cent_deltas_df_study <- full_join(mean_vol_cent_deltas_df, study_info)

abs_vol_lm <- lm(abs_mean_delta_vol_cent ~ study_mean_age + study_n, data=mean_vol_cent_deltas_df_study)
summary(abs_vol_lm)

mean_vol_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_vol_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_vol_lm)[["study_mean_age"]], 
              intercept = coef(abs_vol_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

####Regional Thickness

```{r}
#get subject-level means
mean_ct_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", ct_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_ct_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_ct_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_ct_cent_deltas_df_study <- full_join(mean_ct_cent_deltas_df, study_info)

abs_ct_lm <- lm(abs_mean_delta_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_deltas_df_study)
summary(abs_ct_lm)

mean_ct_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_ct_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_ct_lm)[["study_mean_age"]], 
              intercept = coef(abs_ct_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

much bigger than anticipated based on Nature paper!!? is combatLS overfitting?

```{r}
#get subject-level means
mean_ct_cent_gamlss_df <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct_cent=(mean(centile))*100)
```

```{r}
mean_ct_cent_gamlss_df <- full_join(mean_ct_cent_gamlss_df, study_info)
summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gamlss_df))
summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gamlss_df))

# mean_ct_cent_gamlss_df %>%
#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
#   geom_point(aes(color=study), alpha=0.2) +
#   geom_smooth(method = "lm") +
#   theme(legend.position="none")

#get subject-level means
mean_ct_cent_gam_df <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct_cent=(mean(centile))*100)

mean_ct_cent_gam_df <- full_join(mean_ct_cent_gam_df, study_info)
summary(lm(mean_ct_cent ~ study_mean_age, data=mean_ct_cent_gam_df))
summary(lm(mean_ct_cent ~ study_mean_age + study_n, data=mean_ct_cent_gam_df))

# mean_ct_cent_gam_df %>%
#   ggplot(aes(x=study_mean_age, y=mean_ct_cent)) +
#   geom_point(aes(color=study), alpha=0.2) +
#   geom_smooth(method = "lm") +
#   theme(legend.position="none")
```

####Regional SA

```{r}
#get subject-level means
mean_sa_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", sa_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_sa_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_sa_cent=(mean(abs(centile)))*100) %>%
  dplyr::select(!(starts_with("delta")))

mean_sa_cent_deltas_df_study <- full_join(mean_sa_cent_deltas_df, study_info)

abs_sa_lm <- lm(abs_mean_delta_sa_cent ~ study_mean_age + study_n, data=mean_sa_cent_deltas_df_study)
summary(abs_sa_lm)

mean_sa_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_sa_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_sa_lm)[["study_mean_age"]], 
              intercept = coef(abs_sa_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

####Compare

```{r}
df_list.tmp <- list(mean_sa_cent_deltas_df_study, mean_ct_cent_deltas_df_study, mean_vol_cent_deltas_df_study)
all_reg_deltas <- df_list.tmp %>% reduce(full_join)
```

```{r}
all_reg_deltas_long <- all_reg_deltas %>%
  pivot_longer(cols=starts_with("abs_mean_delta"), names_to="pheno_type", values_to = "abs_cent_delta") %>%
  mutate(pheno_type=gsub("abs_mean_delta_|_cent", "", pheno_type))

all_reg_deltas_long %>%
  ggplot() +
  geom_boxplot(aes(y=abs_cent_delta, fill=pheno_type))
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=study_mean_age_yrs, y=study_mean_delta)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")
delta_sum.df_merged %>%
  ggplot(aes(x=study_mean_age_yrs, y=study_mean_abs.delta)) +
  geom_count(aes(size=study_n)) +
  geom_smooth(method="lm")

age_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
age_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]

age_effects.plt <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=4, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Absolute Difference in Centile Scores") +
  theme(axis.title.x=element_blank())

age_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Pr(>|t|)"]
age_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age_yrs, data=delta_sum.df_merged))$coef["site_mean_age_yrs", "Estimate"]

age_effects.plt.null <- delta_sum.df_merged %>%
  ggplot(aes(x=site_mean_age_yrs, y=(site_mean_delta*100))) +
  geom_count(aes(size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_smooth(method="lm", color="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  annotate("text", x=60, y=-0.3, label=paste("Mean Age \nEstimate =", round(age_effects.plt.slope, 4), "\np =", round(age_effects.plt.p, 4)), size=3) +
  labs(size="site N", x = "Mean Age (Years)", y = "Mean Difference in Centile Scores")

#add prediction to df
delta_sum.df_merged$abs.delta_pred <- predict(full.delta.age.fit)

full.delta.age.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$x,
                                    predicted=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$predicted,
                                    conf.low=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.low,
                                    conf.high=ggeffects::ggemmeans(full.delta.age.fit, terms="site_mean_age_yrs")$conf.high)
full.delta.age.fit.df <- full.delta.age.fit.df %>%
  mutate(predicted = predicted*100,
         conf.low=conf.low*100,
         conf.high=conf.high*100)

marg.age_effects.plt <- ggplot() +
  geom_count(data=delta_sum.df_merged, aes(x=site_mean_age_yrs, y=(abs.delta_pred*100), size=site_n, color=sex_prop_Female)) +
  paletteer::scale_color_paletteer_c("ggthemes::Blue", name = "Proportion \nFemale") +
  geom_line(aes(x=full.delta.age.fit.df$x, y = full.delta.age.fit.df$predicted), size = 0.7, color="#2A5783FF") +
  geom_ribbon(aes(x=full.delta.age.fit.df$x,y= NULL, ymin= full.delta.age.fit.df$conf.low, ymax= full.delta.age.fit.df$conf.high), alpha=0.1, fill="#2A5783FF") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Mean Age (Years)", y = "Predicted Mean Absolute Difference in Centile Scores")
```

are studies w older subjects changing more because there are fewer of them?
check for association between study size and age
```{r}
summary(lm(study_mean_age_yrs ~ study_n, data=delta_sum.df_merged)) #study size is associated with age, but assoc. remains when controlling for size
```

```{r}
cent_deltas_df %>%
  ggplot() +
  geom_histogram(aes(x=age_days)) +
  labs(title="subject distribution across age")

delta_sum.df_merged %>%
  ggplot() +
  geom_histogram(aes(x=study_mean_age_yrs)) +
  labs(title="study distribution across age")
```


##Prop Female

```{r}
ggplot(delta_sum.df_merged) +
  geom_histogram(aes(x=sex_prop_Female))
```

ComBatGAM mean centile ~ % female

```{r}
gam_centile_study.mean.df <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(study) %>%
  summarize(mean_gam_cent=mean(centile),
            mean_gam_abs_cent=mean(abs(centile)))

gam_centile_study.mean.df2 <- left_join(gam_centile_study.mean.df, delta_sum.df_merged)

gam.prop.f.test <- lm(mean_gam_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
summary(gam.prop.f.test)
gam.prop.f.test2 <- lm(mean_gam_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gam_centile_study.mean.df2)
summary(gam.prop.f.test2)

ggplot(gam_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gam_cent)) +
  geom_point() +
  geom_smooth(method="lm")
```

ComBatLS mean centile ~ % female

```{r}
gamlss_centile_study.mean.df <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(study) %>%
  summarize(mean_gamlss_cent=mean(centile),
            mean_gamlss_abs_cent=mean(abs(centile)))

gamlss_centile_study.mean.df2 <- left_join(gamlss_centile_study.mean.df, delta_sum.df_merged)

gamlss.prop.f.test <- lm(mean_gamlss_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
summary(gamlss.prop.f.test)
gamlss.prop.f.test2 <- lm(mean_gamlss_abs_cent ~ sex_prop_Female + study_n + study_mean_age, data=gamlss_centile_study.mean.df2)
summary(gamlss.prop.f.test2)

ggplot(gamlss_centile_study.mean.df2, aes(x=sex_prop_Female, y=mean_gamlss_cent)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
gamlss_centile_site.mean.df3 <- gamlss_centile_site.mean.df2 %>%
  mutate(combat = "gamlss") %>%
  rename(mean_cent = mean_gamlss_cent)
gam_centile_site.mean.df3 <- gam_centile_site.mean.df2 %>%
  mutate(combat = "gam") %>%
  rename(mean_cent = mean_gam_cent)

site_lvl_mean_cents <- full_join(gamlss_centile_site.mean.df3, gam_centile_site.mean.df3)

site_lvl_mean_cents %>%
  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm") +
  ylab("within-site mean centile")
```

```{r}
gamlss_centile_df.mean %>%
  ggplot() + 
  geom_violin(aes(x=sex, y=mean_cent))
gamlss_centile_df.mean$combat <- "gamlss"

gam_centile_df.mean %>%
  ggplot() + 
  geom_violin(aes(x=sex, y=mean_cent))
gam_centile_df.mean$combat <- "gam"

subj.mean.cents <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)

subj.mean.cents %>%
  ggplot() + 
  geom_boxplot(aes(x=sex, y=mean_cent, fill=combat))

#add sex prop
site.info <- delta_sum.df_merged %>%
  dplyr::select(site, sex_prop_Female, site_n)

subj.mean.cents2 <- left_join(subj.mean.cents, site.info)
subj.mean.cents2 %>%
  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=sex)) +
  geom_point(aes(size=site_n)) +
  geom_smooth(method="lm")+
  facet_wrap(~combat)

subj.mean.cents2 %>%
  ggplot(aes(x=sex_prop_Female, y=mean_cent, color=combat, fill=combat)) +
  # geom_point(alpha=0.4) +
  geom_smooth(method="lm") +
  ylab("within-subject mean centile")
```

delta ~ % female

```{r}
# %female
summary(lm(site_mean_abs.delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ sex_prop_Female + site_n, data=delta_sum.df_merged)) #add site size covariate
summary(lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged)) #add site size & age covariates
full.delta.sex.fit <- lm(site_mean_delta ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged) #scaling for plot

sjPlot::plot_model(full.delta.sex.fit, type = "pred", terms = "sex_prop_Female") 
```

```{r}
delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100), color=as.factor(majority), fill=as.factor(majority))) +
  geom_count(aes(shape=as.factor(majority), size=sex_prop), alpha=0.5) +
  geom_line(aes(group=site))
```

```{r}
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="lm")
delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_abs.delta)) +
  geom_count(aes(size=site_n)) +
  geom_smooth(method="gam")

sex_effects.plt.p <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
sex_effects.plt.slope <- summary(lm((site_mean_delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]

sex_effects.plt <- delta_sum.df_merged %>%
  mutate(site_mean_delta=site_mean_delta*100) %>%
  ggplot(aes(x=sex_prop_Female, y=site_mean_delta)) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm",
              color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Mean Difference in Centile Scores") +
  theme(axis.title.y=element_blank()) +
  annotate("text", x=0.25, y=-0.3, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
  guides(size = "none")

sex_effects.plt.p <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Pr(>|t|)"]
sex_effects.plt.slope <- summary(lm((site_mean_abs.delta*100) ~ sex_prop_Female + site_n + site_mean_age, data=delta_sum.df_merged))$coef["sex_prop_Female", "Estimate"]

sex_effects.plt.null <- delta_sum.df_merged %>%
  ggplot(aes(x=sex_prop_Female, y=(site_mean_abs.delta*100))) +
  geom_count(aes(size=site_n, color=site_mean_age_yrs)) +
  geom_smooth(method="lm", color="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N") +
    annotate("text", x=0.25, y=4, label=paste("Proportion Female \nEstimate =", round(sex_effects.plt.slope, 4), "\np =", round(sex_effects.plt.p, 4)), size=3) +
  theme(axis.title=element_blank()) +
  guides(size = "none")

#add prediction to df
delta_sum.df_merged$delta_pred <- predict(full.delta.sex.fit)

full.delta.sex.fit.df <- data.frame(x=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$x,
                                    predicted=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$predicted,
                                    conf.low=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.low,
                                    conf.high=ggeffects::ggemmeans(full.delta.sex.fit, terms="sex_prop_Female")$conf.high)
full.delta.sex.fit.df <- full.delta.sex.fit.df %>%
  mutate(predicted = predicted*100,
         conf.low=conf.low*100,
         conf.high=conf.high*100)

marg.sex_effects.plt <- ggplot() +
  geom_count(data=delta_sum.df_merged, aes(x=sex_prop_Female, y=(delta_pred*100), size=site_n, color=site_mean_age_yrs)) +
  geom_line(aes(x=full.delta.sex.fit.df$x, y = full.delta.sex.fit.df$predicted), size = 0.7, color="#24693DFF") +
  geom_ribbon(aes(x=full.delta.sex.fit.df$x,y= NULL, ymin= full.delta.sex.fit.df$conf.low, ymax= full.delta.sex.fit.df$conf.high), alpha=0.1, fill="#24693DFF") +
  paletteer::scale_color_paletteer_c("ggthemes::Green", name = "Mean Age \n(Years)") +
  theme_linedraw(base_size = 11) +
  labs(size="site N", x = "Proportion Female", y = "Predicted Mean Difference in Centile Scores")

delta_sum.by.sex.df %>%
  ggplot(aes(x=sex, y=(mean_delta*100))) +
  geom_boxplot()
```

```{r}
# age*%female
summary(lm(site_mean_abs.delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
summary(lm(site_mean_delta ~ site_mean_age_yrs*sex_prop_Female, data=delta_sum.df_merged))
```

```{r}
cent_deltas_df %>%
  ggplot(aes(x=age_days, y=(delta.GMV*100))) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")

cent_deltas_df %>%
  ggplot(aes(x=delta.lh_SA_fusiform*100, color=sex, fill=sex)) +
  geom_histogram(aes(y=after_stat(c(
      count[group==1]/sum(count[group==1]),
      count[group==2]/sum(count[group==2])
    )*100)), alpha=0.4, position="identity", bins=60) +
  ylab("Percentage")

cent_deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_supramarginal*100, color=sex, fill=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth(method="lm")

cent_deltas_df %>%
  ggplot(aes(x=age_days, y=delta.rh_Vol_temporalpole*100, color=sex, fill=sex)) +
  geom_point(alpha=0.4) +
  geom_smooth(method="lm")
```

#FIGURE 5

```{r}
# study demographics
fig5 <- ggarrange(
ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          # align = "hv", 
          common.legend = TRUE,
          legend = "bottom",
          widths=c(2, 1),
          heights=c(14, 1)),
# mean effects
ggarrange(age_effects.plt,
sex_effects.plt,
nrow=2, ncol=1,
align="hv",
labels=c("B", "C")),
common.legend=FALSE,
nrow=1, ncol=2,
labels=c("A", NA),
widths=c(1, 1.4)
)

ggsave("figs/figure_5.jpeg", fig5, width=12, height=10, units="in")
```

#ComBatted Data

```{r}
gam_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gam_batch.study_data.csv"))
gam_data_df <- gam_data_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)

gamlss_data_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gamlss_batch.study_data.csv"))
gamlss_data_df <- gamlss_data_df %>% mutate(study=sub("\\_.*", "", site),
         age_years=age_days/365.25)
```

```{r}
gam_data_df %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
  theme(legend.position="none")

gamlss_data_df %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh_CT_medialorbitofrontal, color=study), alpha=0.4) +
  theme(legend.position="none")
```

```{r}
#get subject-level means
gam_data_ct_avg <- gam_data_df %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct=mean(val)) #convert units to centiles space

gamlss_data_ct_avg <- gamlss_data_df %>%
  pivot_longer(cols=any_of(ct_list), names_to="pheno", values_to = "val") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_ct=mean(val)) #convert units to centiles space
```

```{r}
gam_data_ct_avg %>%
  ggplot() +
  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
  theme(legend.position="none")

gamlss_data_ct_avg %>%
  ggplot() +
  geom_point(aes(x=age_days, y=mean_ct, color=study), alpha=0.4) +
  theme(legend.position="none")
```

```{r}
gam_data_ct_avg$combat <- "gam"
gamlss_data_ct_avg$combat <- "gamlss"

df_list.tmp <- list(gam_data_ct_avg, gamlss_data_ct_avg, study_info)
ct_avg_data <- df_list.tmp %>% reduce(full_join)
```

```{r}
ct_avg_data %>%
  ggplot() +
  geom_point(aes(x=age_days, y=mean_ct, color=combat), alpha=0.4)

ct_avg_data %>%
  ggplot() +
  geom_point(aes(x=study_mean_age, y=mean_ct, color=combat), alpha=0.2)

ct_avg_data %>%
  ggplot() +
  geom_point(aes(x=study_mean_age, y=mean_ct, color=study), alpha=0.4) +
  facet_wrap(~combat) +
  theme(legend.position="none")


```
