---
title: "hbm_paper_results"
output: html_document
date: '2023-11-12'
---

Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")
```
```{r}
source("./cubic_scripts/R_scripts/stats_tests.R")
# devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
```


```{r}
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives"

#pheno lists
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

#diff lists
pheno_diff_list <- paste0("diff_", pheno_list)
pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)


#z lists
pheno_list.z <- paste0(pheno_list, ".z")
z.pheno_diff_list <- paste0(pheno_diff_list, ".z")
z.pheno_abs.diff.list <- paste0(pheno_abs.diff.list, ".z")

#also load raw data
ukb_cn_df <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/ukb_to_model/ukb_CN_data_simsites.csv")

#years for plotting
ukb_cn_df <- ukb_cn_df %>%
  mutate(age_years = age_days/365.25)

#write out new dk atlas w/o medial wall
dk_nomed <- dk
dk_nomed$data <- dk_nomed$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")
#plotting info from dk atlas
dk.parc <- dk_nomed$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()
```

# Main Analyses

## Load Data

pulled variance from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_variance.csv' .`

Get predicted centiles
- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_predictions.csv' .`

```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
  
  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data)
}
```

**QC: find ppts with centiles at 1 or 0**

```{r}
combined_preds %>% 
  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
  dplyr::select(participant) %>%
  unique()

combined_preds %>% 
  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
  dplyr::select(participant) %>%
  unique()
```

saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!

Get errors
- pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_diffs.csv' .`

```{r}
raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_diffs.csv", full.names = TRUE) %>% unlist()

diffs.df <- fread(file = raw_files, sep=",") %>%
     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
```

Get subject-wise summaries
- pulled from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/subject-wise/*.csv' .`

```{r}
raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_subj_pred.csv", full.names = TRUE) %>% unlist()

subj_sum.df <- fread(file = raw_files, sep=",") %>%
     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))
```

Get differences in male vs female predicted variances at mean sample age

```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File),
           sex_standard = sex_effect/f.var)
  
  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset),
         dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)

sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")

plot.info <- sigma.sex.df %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
  unique() %>%
  mutate(dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset))

combined_var <- inner_join(combined_var, plot.info)
```

## SCALE EFFECTS OF SEX

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
#var.plot

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
```

```{r}
#plot global
var.plot.global.df <- combined_var %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
#var.plot.global
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))
p_annotate
ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```


## CENTILES

Do centile scores change based on combat version?
Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)

Within each feature (phenotype), iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

```{r, eval=FALSE}
full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
```

```{r}
full_cent_res <- readRDS("./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
full_cent_res %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            min_pval=min(p.val_fdr),
            max_pval=max(p.val_fdr)) %>%
  dplyr::filter(!grepl('_refA', comp))#drop refA levels

full_cent_res %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  # dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
  ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r, eval=FALSE}
full_cent_res%>%
  dplyr::filter(comp =="cf.gamlss_cf" & bigger_group == "cf.gamlss" & sig_fdr ==TRUE)

diffs.df %>%
  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
  ggplot() +
  geom_boxplot(aes(y=abs.diff_rh_CT_parstriangularis, fill=dataset))

diffs.df %>%
  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
  ggplot() +
  geom_boxplot(aes(y=diff_rh_CT_parstriangularis, fill=dataset))
```

Little plot
```{r}
#get means w/in feature 
mean.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno_abs.diff.list) %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
```


```{r, eval=FALSE}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
cent.diffs.plot

ggsave("figs/win_feature_mean_cent_abs.diffs.jpeg", cent.diffs.plot)
```

Plot errors for global phenotypes

get stat results for global phenos:
```{r, message=FALSE}
full_cent_res %>%
  dplyr::filter(pheno %in% vol_list_global) %>%
  dplyr::filter(!grepl('_refA', comp))

full_cent_res %>%
  dplyr::filter(bigger_group == "cf.gamlss")%>%
  dplyr::filter(!grepl('_refA', comp))
```

all significant

```{r}
global_abs.diffs <- diffs.df %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", vol_list_global)) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")

global_cent_errs_plot <- global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
         abs.diff = abs.diff*100) %>% #for visuals
  ggplot(aes(pheno, abs.diff)) +
  geom_violin(aes(x=pheno, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #WMV
geom_signif(y_position = c(29, 30.5, 4), xmin = c(0.7, 0.7, 1), 
              xmax = c(1, 1.3, 1.3), annotation = c("***","***", "***"),
              tip_length = 0, vjust = 0.6) +
  #GMV
  geom_signif(y_position = c(30, 31.5, 3), xmin = c(1.7, 1.7, 2), 
              xmax = c(2, 2.3, 2.3), annotation = c("***","***", "***"),
              tip_length = 0, vjust = 0.6) +
  #sGMV
  geom_signif(y_position = c(28, 29.5, 2.8), xmin = c(2.7, 2.7, 3), 
              xmax = c(3, 3.3, 3.3), annotation = c("***","***", "***"),
              tip_length = 0, vjust = 0.6) +
  #Ventricles
  geom_signif(y_position = c(24, 25.5, 7.5), xmin = c(3.7, 3.7, 4), 
              xmax = c(4, 4.3, 4.3), annotation = c("***","***", "***"),
              tip_length = 0, vjust = 0.6) +
  labs(y="Absolulte Centile Error", x = "Feature") +
  theme(plot.title = element_text(size=12))
global_cent_errs_plot
# ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
```

plot median on brain
```{r}
#get median w/in feature 
med.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno_abs.diff.list) %>%
  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), median)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="median_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

med.diffs.long2 <- left_join(med.diffs.long, dk.parc, by="label")
max(med.diffs.long2$median_centile_abs.diff)*100

dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
med.diffs.plot.df <- med.diffs.long2 %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         median_centile_abs.diff = median_centile_abs.diff*100)

med.diffs.plot <- ggplot(med.diffs.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_centile_abs.diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5)) #
# med.diffs.plot
```

```{r}
#plot global
med.diffs.plot.global.df <- med.diffs.long2 %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         median_centile_abs.diff = median_centile_abs.diff*100)

med.diffs.plot.global <- ggplot(med.diffs.plot.global.df) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_centile_abs.diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Oslo", name = "Median Absolute Centile Error", limits = c(0, 7.5))
# med.diffs.plot.global
```

Combine plots - need to keep playing around to make the spacing pretty when it saves out
```{r}
p_diff <- ggarrange(
  med.diffs.plot.global, med.diffs.plot,
  ncol=2, nrow=1, 
  widths = c(.95, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

# p_diff_annotate <- annotate_figure(p_diff)
# p_diff_annotate
```
plot w ref sites
```{r, eval=FALSE}
#get means w/in feature 
mean.cent.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="cent_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  # mutate(pheno=gsub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

mean.cent.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=cent_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_grid(dataset~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Absolute errors in centiles of brain structure features across ComBat processing pipelines")
```
Quick comp of cf.gam and cf.gamlss

```{r}
gam.gamlss.diffs <- diffs.df %>%
  dplyr::filter(dataset=="cf.gam" | dataset=="cf.gamlss")

gam_better <- gam.gamlss.diffs %>%
  group_by(dataset) %>%
  summarize_at(pheno_abs.diff.list, median) %>%
  ungroup() %>%
  pivot_longer(cols = all_of(pheno_abs.diff.list), names_to="pheno") %>%
  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
         pheno = sub("abs.diff_", "", pheno)) %>%
  ungroup() %>%
  dplyr::filter(gam_minus_gamlss < 0)

gam_better$pheno

ns_gam_gamlss_comp <- full_cent_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)

intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
```
```{r, eval=FALSE}
for (pheno in pheno_abs.diff.list){
  p <- ggplot(gam.gamlss.diffs, aes(y = gam.gamlss.diffs[[pheno]], fill= dataset)) +
      geom_boxplot() +
    ggtitle(pheno)
print(p)
}
```

**21 regions where ComBat GAM is better than ComBat-LS! All cortical thickness -> 206-21 = 185 regions where ComBat-LS fits better.

### Sex-Biases

Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)

```{r, eval=FALSE}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  facet_wrap(~dataset) +
  theme(legend.position="none") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")

subj_sum.df %>%
  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
  geom_violin(trim=FALSE, scale="width") +
  labs(y="Centile Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean error in centile scores across ComBat configurations")

subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
  geom_violin(trim=FALSE) +
  facet_wrap(~dataset) +
  labs(y="Magnitude Centile Error", x = "Simulated Site") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")

#look at sex distributions of ppts across sites
subj_sum.df %>%
  ggplot() +
  geom_bar(aes(x=sim.site, fill=sex))+
  ggtitle("Distributions of male and female subjects across simulated sites")
```

Significance testing (w/in sub. mean abs centile error)  - using wilcox (mann whitney u) since distributions are skewed but variances are ~= across sexes

```{r}
sex.bias.wilcox.tests(subj_sum.df)
sex.bias.wilcox.tests(subj_sum.df, to_test="mean_cent_diff")
```

Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features). However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.

Testing for significant M:F differences in centile erros WITHIN each feature!

check for skew (welch t.test vs ranked welch t test)
```{r, eval=FALSE}
for (pheno in pheno_list){
  val <- paste0("diff_", pheno)
  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
      geom_histogram(alpha=0.5, position="identity") +
    ggtitle(val) +
    facet_wrap(~dataset, scales="free")
print(p)
}
```

```{r, eval=FALSE}
for (cf in c("cf.gam", "cf.gamlss")) {
  for (pheno in vol_list_global){
  val <- paste0("diff_", pheno)
  med.df <- diffs.df %>%
    dplyr::filter(dataset == cf) %>%
    group_by(sex) %>%
    summarize_at(val, median, na.rm=FALSE)

 p <- diffs.df %>%
    dplyr::filter(dataset == cf) %>%
    ggplot(aes(x=.[[val]], fill = sex, color=sex)) +
      geom_histogram(alpha=0.5, position="identity") +
    ggtitle(paste(val, cf)) +
    geom_vline(aes(xintercept = med.df[med.df$sex=="Female", ][[val]])) +
    geom_vline(aes(xintercept = med.df[med.df$sex=="Male", ][[val]]))
  print(p)
  }
}
```

not normally distributed, use nonparametric t test of ranks with welch correction for different variances. Also calculating the median centile errors of males and females, since this seems like a more intuitive thing to plot than effect size (since the effect size is in units of rank, not actually centiles)

```{r, message=FALSE, eval=FALSE}
sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=pheno_diff_list)
saveRDS(sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_out.RDS")
```

```{r}
sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_cent_sexbias_all_out.RDS")
#also get summary
sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
```

```{r,eval=FALSE}
#view combat gam and combat LS ns
ns.sex_diff_filt.gam <- sex_t_tests_in_feat.df %>%
  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gam") %>%
  arrange(pheno)
ns.sex_diff_filt.gamlss <- sex_t_tests_in_feat.df %>%
  dplyr::filter(sig_fdr==FALSE & dataset =="cf.gamlss") %>%
  arrange(pheno)
```

quick plot for median diffs
```{r}
sex_t_tests_in_feat.df %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>%
  ggplot() +
  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

sex_t_tests_in_feat.df %>%
  group_by(dataset) %>%
  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
```

```{r}
plt.df <- diffs.df %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>% #select centile errors only
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         diff = diff*100,
         pheno_cat = factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
```

stat test re-do w wilcox 
```{r}
stat.test <- plt.df %>%
  group_by(dataset, pheno) %>%
  wilcox_test(diff ~ sex) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")
```
```{r}
stat.test2 <- stat.test %>%
  add_xy_position(x = "pheno", dodge = 0.8, fun="max") %>%
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*")) #simplify significance labels
```

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")

big_violin_plt <- plt.df %>%
  ggplot(aes(x=pheno, y=diff)) +
  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
  # geom_violin(aes(fill=sex, color=sex)) +
  theme_linedraw(base_size = 11) +
  labs(y="Centile Error", x = "Brain Feature") +
  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
  labs(fill = "Sex", color="Sex")
  # ggtitle("Male and female centile errors")
big_violin_plt
ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
```

```{r, eval=FALSE}
big_violin_plt_wlabels <- plt.df %>%
  ggplot(aes(x=pheno, y=diff)) +
  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
  # geom_violin(aes(fill=sex, color=sex)) +
  theme_linedraw(base_size = 11) +
  labs(y="Centile Error", x = "Brain Feature") +
  theme(plot.title = element_text(size=12)) +
  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 2.5, tip.length = 0, remove.bracket = TRUE) +
  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
  labs(fill = "Sex", color="Sex")
  # ggtitle("Male and female centile errors")
big_violin_plt_wlabels
ggsave("figs/big_sex_violin_plt_wlabels.jpeg", big_violin_plt_wlabels, width=16, height=9, units="in")
```


```{r}
sex_t_tests_in_feat.df2 <- sex_t_tests_in_feat.df %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    label = sub("_[^_]*_", "_", pheno)) #for plotting
sex_t_tests_to_plot <- left_join(sex_t_tests_in_feat.df2, dk.parc, by="label")

# max(sex_t_tests_to_plot$median_sex_diff)*100
# min(sex_t_tests_to_plot$median_sex_diff)*100
```


plot w/in feature t tests

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
sex.var.plot.df <- sex_t_tests_to_plot %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                                     TRUE ~ median_sex_diff*100))

sex.var.plot <- ggplot(sex.var.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  #scale_fill_gradient2() +
  # ggseg::geom_brain(data=med_wall, atlas = dk, 
  #            position = position_brain(hemi ~ side),
  #            aes(fill=median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-6, 6)) #
sex.var.plot
```

```{r}
#plot global
sex.var.plot.global.df <- sex_t_tests_to_plot %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(dataset %in% c("cf.lm", "cf.gam", "cf.gamlss")) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
        median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>% #drop NS values
  group_by(dataset)

tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork",  limits = c(-6, 6), name = "Male - Female Median \nCentile Error") #,
# sex.var.plot.global
```

Combine plots - need to keep playing around to make the spacing pretty when it saves out
```{r}
p_sex <- ggarrange(
  sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
```

Sex Effect is the median male centile error - median female centile error (i.e. units are in centiles). Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)? Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.


### W/o extremes
Are extreme centiles (<5% and >95%) driving observed results?

Get raw centiles and merge with centile differences

```{r, message=FALSE}
raw_centiles <- combined_preds %>%
  dplyr::filter(dataset == "ukb_CN")

merge.df <- full_join(raw_centiles, diffs.df)
unique(merge.df$dataset)
dim(raw_centiles)
dim(merge.df)

merge.df <- merge.df %>%
  dplyr::select(!File_ID)
```


```{r}
df_list <- c()
count <- 1
for(pheno in pheno_list) {
  list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))

  df <- merge.df %>%
    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
    group_by(participant) %>% 
    dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "ukb_CN")) %>% #drop ppts if raw centile is <0.05 or >0.95
    ungroup() %>%
    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
  df_list[count] <- list(df)
  count <- count + 1
}
length(df_list) == length(pheno_list)
```
merge into single dataframe
```{r, message=FALSE}
no_ext <- Reduce(full_join, df_list)

#make sure no cols are lost
dim(no_ext)
# #make sure no cols are lost
# ncol(no_ext) == ncol(merge.df) #lost File_ID col, not going to bother rerunning rn
# nrow(no_ext)
```

Rerun stats tests

centile error diffs
```{r, eval=FALSE}
noext_cent_res <- centile.t.tests.full_result(no_ext, feature_list=pheno_abs.diff.list)
saveRDS(noext_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_noext_out.RDS")
```

```{r}
noext_cent_res <- readRDS("./derivatives/ukb_basic/featurewise_cent_t_tests_all_noext_out.RDS")
noext_cent_res %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            min_pval=min(p.val_fdr),
            max_pval=max(p.val_fdr)) %>%
  dplyr::filter(!grepl('_refA', comp))#drop refA levels

noext_cent_res %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
  ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Little plot
```{r, eval=TRUE}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

#get means w/in feature 
no.ext_mean.diffs.long <- no_ext %>%
  dplyr::select(sex, dataset, pheno, pheno_abs.diff.list) %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean, na.rm=TRUE)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
```
```{r, eval=FALSE}
no.ext_mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("No extremes: Absolute errors in centile scores of brain structure features across ComBat processing pipelines")
cent.diffs.plot
```

```{r}
no.ext_mean.diffs.long$type <- "no_ext"
mean.diffs.long$type <- "full"

no.ext_merge <- full_join(no.ext_mean.diffs.long, mean.diffs.long)

no.ext_merge %>%
  ggplot() +
  geom_boxplot(aes(x=type, y=centile_abs.diff, fill=dataset)) +
  ggtitle("W/in feature average magnitude of centile error, w/ and w/o extreme subjects")
```

w/in feature sex-biases
```{r, message=FALSE, eval=FALSE}
noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
saveRDS(noext.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")
```

```{r}
noext.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")

#compare summaries
noext.sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()

sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
```

plot comparative differences in medians
```{r}
sex.var.plot.full.df <- sex_t_tests_to_plot %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>%
  group_by(dataset, pheno_cat) %>%
  dplyr::select(dataset, pheno_cat, pheno, median_sex_diff)

sex.var.plot.full.df$type <- "full"

noext.sex_t_tests_in_feat.df2 <- noext.sex_t_tests_in_feat.df %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    label = sub("_[^_]*_", "_", pheno)) %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>%
  dplyr::select(dataset, pheno_cat, pheno, median_sex_diff) %>%
  group_by(dataset, pheno_cat)

noext.sex_t_tests_in_feat.df2$type <- "no_ext"

#merge
noext_plot.df <- full_join(sex.var.plot.full.df, noext.sex_t_tests_in_feat.df2)
```

```{r}
noext_plot.df %>%
  ggplot() +
  geom_col(aes(x=pheno, y=median_sex_diff, fill=type), alpha=0.7, position="identity") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  facet_wrap(~dataset, scale="free_y")

noext_plot.df %>%
  ggplot() +
  geom_violin(aes(x=type, y=median_sex_diff, fill=type), trim=FALSE) +
  facet_wrap(~dataset, scales="free_y")
```

## Z-SCORES

Do z scores change based on combat version?

check for skew (welch t.test vs ranked welch t test)
```{r, eval=FALSE}
for (val in z.pheno_abs.diff.list){
  p <- ggplot(diffs.df, aes(x = diffs.df[[val]])) +
      geom_histogram(alpha=0.5, position="identity") +
    ggtitle(val) +
    facet_wrap(~dataset, scales="free")
print(p)
}
```

Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)

Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

```{r, eval=FALSE}
full_z_res <- centile.t.tests.full_result(diffs.df, feature_list=z.pheno_abs.diff.list)
saveRDS(full_z_res, "./derivatives/ukb_basic/featurewise_z_t_tests_all_out.RDS")
```

```{r}
full_z_res <- readRDS("./derivatives/ukb_basic/featurewise_z_t_tests_all_out.RDS")
full_z_res %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            min_pval=min(p.val_fdr),
            max_pval=max(p.val_fdr)) %>%
  dplyr::filter(!grepl('_refA', comp))#drop refA levels

full_z_res %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  # dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
  ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Little plot
```{r}
#get means w/in feature 
mean.z.diffs.long <- diffs.df %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno=gsub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

z.diffs.plot <- mean.z.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
z.diffs.plot

ggsave("figs/win_feature_mean_z_abs.diffs.jpeg", z.diffs.plot)
```

plot w ref sites
```{r, eval=FALSE}
#get means w/in feature 
mean.z.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno=gsub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

mean.z.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_grid(dataset~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
```

Quick comp of cf.gam and cf.gamlss

```{r}
gam_better.z <- gam.gamlss.diffs %>%
  group_by(dataset) %>%
  summarize_at(z.pheno_abs.diff.list, median) %>%
  ungroup() %>%
  pivot_longer(cols = all_of(z.pheno_abs.diff.list), names_to="pheno") %>%
  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
         pheno = sub("abs.diff_", "", pheno)) %>%
  ungroup() %>%
  dplyr::filter(gam_minus_gamlss < 0)

gam_better.z$pheno

ns_gam_gamlss_comp.z <- full_z_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)

intersect(gam_better.z$pheno, ns_gam_gamlss_comp.z$pheno)
```

### Sex-Biases

Visualize

```{r, eval=FALSE}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot() +
  geom_boxplot(aes(x=sex, y=mean_z_abs.diff, fill=sex)) +
  labs(y= "Magnitude Z Error", x = "ComBat Configurations") +
  facet_wrap(~dataset, scales="free_y") +
  theme(legend.position="none") +
  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat configurations")

subj_sum.df %>%
  ggplot(aes(x=dataset, y=mean_z_diff, fill=sex)) +
  geom_violin(trim=FALSE, scale="width") +
  labs(y="Z Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean error in z-scores across ComBat configurations")

subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
  geom_violin(trim=FALSE) +
  facet_wrap(~dataset) +
  labs(y="Magnitude Z Error", x = "Simulated Site") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude error in z-scores across \nComBat configurations")

#look at sex distributions of ppts across sites
subj_sum.df %>%
  ggplot() +
  geom_bar(aes(x=sim.site, fill=sex), position="fill")+
  ggtitle("Distributions of male and female subjects across simulated sites")
```

Significance testing (w/in sub. mean abs z error) 

```{r}
sex.bias.t.tests(subj_sum.df, to_test="mean_z_diff")
sex.bias.t.tests(subj_sum.df, to_test="mean_z_abs.diff")
```

Testing for significant M-F differences in z score errors WITHIN each feature!

check for skew (welch t.test vs ranked welch t test)
```{r, eval=FALSE}
for (val in z.pheno_diff_list){
  p <- ggplot(diffs.df, aes(x = diffs.df[[val]], fill= sex, color=sex)) +
      geom_histogram(alpha=0.5, position="identity") +
    ggtitle(val) +
    facet_wrap(~dataset, scales="free")
print(p)
}
```

skewed but ~ variances? sticking with t test of ranks. Also calculating the median z-score errors of males and females

```{r, message=FALSE, eval=FALSE}
z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(diffs.df, feature_list=z.pheno_diff_list)
saveRDS(z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_out.RDS")
```

```{r}
z.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_z_sexbias_all_out.RDS")

#also get summary
z.sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
```

### W/o extremes
Are extreme z-scores (<-2 and >2) driving observed results?

Get raw centiles and merge with centile differences

```{r}
z.df_list <- c()
count <- 1
for(pheno in pheno_list.z) {
  list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))

  df <- merge.df %>%
    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
    group_by(participant) %>% 
    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "ukb_CN")) %>% #drop ppts if raw z score is < -2 or > 2
    ungroup() %>%
    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
  z.df_list[count] <- list(df)
  count <- count + 1
}
length(z.df_list) == length(pheno_list.z)
```
merge into single dataframe
```{r, message=FALSE}
z.no_ext <- Reduce(full_join, z.df_list)

#make sure no cols are lost
dim(z.no_ext)
```

Rerun stats tests

z error diffs
```{r, eval=FALSE}
noext_z_res <- centile.t.tests.full_result(z.no_ext, feature_list=z.pheno_abs.diff.list)
saveRDS(noext_z_res, "./derivatives/ukb_basic/featurewise_z_t_tests_all_noext_out.RDS")
```

```{r}
noext_z_res <- readRDS("./derivatives/ukb_basic/featurewise_z_t_tests_all_noext_out.RDS")
noext_z_res %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig_regions = sum(sig_fdr),
            min_pval=min(p.val_fdr),
            max_pval=max(p.val_fdr)) %>%
  dplyr::filter(!grepl('_refA', comp))#drop refA levels

noext_z_res %>%
  dplyr::filter(sig_fdr==TRUE) %>%
  dplyr::filter(!grepl('_refA', comp)) %>%#drop refA levels
  ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Little plot
```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

#get means w/in feature 
z.no.ext_mean.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean, na.rm=TRUE)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
```

```{r}
z.no.ext_mean.diffs.long$type <- "no_ext"
mean.z.diffs.long$type <- "full"

z.no.ext_merge <- full_join(z.no.ext_mean.diffs.long, mean.z.diffs.long)

z.no.ext_merge %>%
  ggplot() +
  geom_boxplot(aes(x=type, y=z_abs.diff, fill=dataset)) +
  ggtitle("W/in feature average magnitude of z error, w/ and w/o extreme subjects")
```

w/in feature sex-biases
```{r, message=FALSE}
noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)
saveRDS(noext.z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
```

```{r}
noext.z.sex_t_tests_in_feat.df <- readRDS("./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")

#compare summaries
noext.z.sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()

z.sex_t_tests_in_feat.df %>%
    group_by(dataset) %>%
    dplyr::summarize(n_sig_regions = sum(sig_fdr),
              min_pval = min(p.val_fdr),
              max_pval = max(p.val_fdr)) %>%
    ungroup() %>%
    as.data.frame()
```

## CENTILE FANS

pulled predicted centiles from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/centile_csvs/*_centiles.csv' .`

```{r}
raw_files <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_centiles.csv", full.names = TRUE)

df_list <- list() #new empty list
for (file in raw_files) {
  print(paste("reading file", file))
  # Read each CSV file
    data <- fread(file)
    
    data <- data %>%
     mutate(dataset = gsub("-data", "", dataset))

    # Bind the data to the combined dataframe
    df_list[[file]] <- as.data.frame(data)
    assign("basic_centiles", df_list)
}
# length(basic_centiles)
centiles.df <- bind_rows(basic_centiles)
#skimr::skim(centiles.df)
# names(centiles.df)
```

just work with global volumes to visualize for now

```{r}
#get raw data points for visualization
global.df.long <- ukb_cn_df %>%
  dplyr::select(participant, age_days, sex, vol_list_global) %>%
  pivot_longer(cols = vol_list_global, values_to = "value", names_to="pheno")
```


```{r, eval=FALSE}
global.cent.df <- centiles.df %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('-refA', dataset))


for (feature in unique(global.cent.df$pheno)) {

  df <- global.cent.df %>%
    dplyr::filter(pheno == feature)
  
 plot <- ggplot() +
      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
      labs(title=paste("Male", feature, "brain chart"))
 
 print(plot)
 
 plot <- ggplot() +
      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
      geom_line(aes(x=df$age_days, y=df$F_centile_1, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_2, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_3, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_4, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_5, color=df$dataset), linewidth=1.5) +
      geom_line(aes(x=df$age_days, y=df$F_centile_6, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_7, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_8, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$F_centile_9, color=df$dataset)) +
      labs(title=paste("Female", feature, "brain chart"))
 
 print(plot)
 
 plot <- ggplot() +
      geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
      geom_line(aes(x=df$age_days, y=df$centile_1, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_2, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_3, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_4, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_5, color=df$dataset), linewidth=1.5) +
      geom_line(aes(x=df$age_days, y=df$centile_6, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_7, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_8, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$centile_9, color=df$dataset)) +
      labs(title=paste(feature, "brain chart"))
 
 print(plot)
  
}
```

Looking at WMV more specifically for visualizations
```{r}
#wm centile lines
wmv.df <- centiles.df %>%
  dplyr::filter(pheno == "WMV") %>%
  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
  #pivot_longer(cols=contains("centile"), name)  %>%
  pivot_longer(
    cols = contains("centile"),
    names_to = c("sex", ".value"),
    names_pattern = "([MF]?)_(centile_\\d+)",
    values_to = "value"
  ) %>%
 dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
                                   sex == "F" ~ "Female")),
         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
         age_years = age_days/365.25,
         label = ifelse(age_days == max(age_days), "Test", NA))
```

adding jitter to age (x-axis) because otherwise frankly it's hard to look at
centiles = c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)

```{r}
dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")

tmp_labels <- wmv.df %>%
  dplyr::filter(age_days == max(age_days) & dataset=="ukb-CN-simsites") %>%
  pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
  mutate(centile = case_when(centile_num == 1 ~ "1%", centile_num==2 ~ "5%", centile_num==3 ~ "10%", 
                             centile_num == 4 ~ "25%", centile_num==5 ~ "50%", centile_num==6 ~ "75%",
                             centile_num == 7 ~ "90%", centile_num==8 ~ "95%", centile_num==9 ~ "99%")) %>%
  dplyr::filter(centile_num %in% c(1, 2, 5, 8, 9))

wmv_centile_fan <- ggplot(wmv.df) +
  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$WMV), alpha=0.4, width=.2, color="lightgray") +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  #geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~sex) + 
  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile)), nudge_x = 0.35, size=2.5) +
  xlab("Age (yrs)") + ylab("White Matter Volume (mm3)") +
  theme_linedraw(base_size = 11) +
  theme(plot.title = element_text(size=12), legend.position="bottom")
wmv_centile_fan
```
Look at other phenotypes

```{r}
#looking at other phenotypes
ven.df <- centiles.df %>%
  dplyr::filter(pheno == "Ventricles") %>%
  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss"| dataset == "cf.lm") %>%
  #pivot_longer(cols=contains("centile"), name)  %>%
  pivot_longer(
    cols = contains("centile"),
    names_to = c("sex", ".value"),
    names_pattern = "([MF]?)_(centile_\\d+)",
    values_to = "value"
  ) %>%
 dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
                                   sex == "F" ~ "Female")),
         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ),
         age_years = age_days/365.25)
tmp_labels <- ven.df %>%
  dplyr::filter(age_years == max(age_years) & dataset=="ukb-CN-simsites") %>%
  pivot_longer(cols=starts_with("centile"), values_to = "value", names_to="centile_num", names_prefix = "centile_") %>%
  mutate(centile = case_when(centile_num == 1 ~ "1%", centile_num==2 ~ "5%", centile_num==3 ~ "10%", 
                             centile_num == 4 ~ "25%", centile_num==5 ~ "50%", centile_num==6 ~ "75%",
                             centile_num == 7 ~ "90%", centile_num==8 ~ "95%", centile_num==9 ~ "99%")) %>%
  dplyr::filter(centile_num %in% c(1, 2, 5, 8, 9)) %>%
  #repel for visuals
  mutate(value = ifelse(centile_num == 1, value*.75,
                           value))

ven_plot <- ggplot(ven.df) +
  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$Ventricles),  alpha=0.4, width=.2, color="lightgray") +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~sex) + 
  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
  theme_linedraw(base_size = 11) +
  geom_label(data=tmp_labels, aes(x=age_years, y=value, label=as.character(centile))) +
  xlab("Age (days)") + ylab("Volume (mm3)") +
  labs(title="Ventricular volume brain charts")
```

```{r}
ven.df %>%
  dplyr::filter(sex=="Female") %>%
ggplot() +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset_names)) + 
  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
  xlab("Age (yrs)") + ylab("Volume (mm3)") +
  theme_classic(base_size = 11) +
  theme(plot.title = element_text(size=12), legend.position="none")
```

Random cortical feature

```{r}
r.sa.pc.df <- centiles.df %>%
  dplyr::filter(pheno == "rh_SA_postcentral") %>%
  dplyr::filter(dataset == "ukb-CN-simsites" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  #pivot_longer(cols=contains("centile"), name)  %>%
  pivot_longer(
    cols = contains("centile"),
    names_to = c("sex", ".value"),
    names_pattern = "([MF]?)_(centile_\\d+)",
    values_to = "value"
  ) %>%
 dplyr::filter(!if_all(c(sex,contains("centile")), is.na)) %>%
  mutate(sex = as.factor(case_when(sex == "M" ~ "Male",
                                   sex == "F" ~ "Female")),
         dataset = fct_relevel(dataset, "ukb-CN-simsites", after = Inf ))

ggplot(r.sa.pc.df) +
  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_days, y=ukb_cn_df$rh_SA_postcentral), alpha=0.7, width=150, color="lightgray") +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  geom_line(aes(x=age_days, y=centile_1, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_2, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_3, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
  geom_line(aes(x=age_days, y=centile_6, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_7, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_8, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_days, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~sex) + 
  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
  theme_minimal() +
  xlab("Age (days)") + ylab("Volume (mm3)") +
  theme_minimal() +
  labs(title="SA of right postcentral brain charts")
```

## FIGURE 1: METHODS

toy brain chart for figs, diagram portion done in canva

```{r}
ven.df %>%
  dplyr::filter(sex=="Female") %>%
  mutate(dataset = factor(dataset, levels = c("ukb-CN-simsites", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
ggplot() +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  # geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_3, color=dataset)) +
  # geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset), linewidth=1.5) +
  # geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_7, color=dataset)) +
  # geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  # geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~dataset, nrow=1, labeller=labeller(dataset=dataset_names)) + 
  # scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
  xlab("Age (yrs)") + ylab("Brain Feature") +
  theme_classic(base_size = 11) +
  theme(plot.title = element_text(size=12), legend.position="none", axis.text.y=element_blank())
```

toy error distributions
```{r}
cf.df <- data.frame("err" = c(rnorm(100, -3, 10)),
          "df" = c(rep("cf.lm", 100)))
cf.gam.df <- data.frame("err" = c(rnorm(100, 1, 5)),
          "df" = c(rep("cf.gam", 100)))
cf.gamlss.df <- data.frame("err" = c(rnorm(100, 0, 3)),
          "df" = c(rep("cf.gamlss", 100)))

rbind(cf.df, cf.gam.df, cf.gamlss.df) %>%
  ggplot() +
  geom_histogram(aes(x=err, fill=df), color="black", alpha=0.7, position="identity") +
  scale_fill_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme_classic(base_size=20) +
  xlab("Centile Error") + 
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.line.y = element_blank()) +
  scale_x_continuous(labels=c('', '', '', 0, '', '', ''))
```

## FIGURE 2

not sure whether to include
```{r, eval=FALSE}
#look at sex distributions of ppts across sites
data.p <- subj_sum.df %>%
  count(sim.site, sex) %>% 
  group_by(sim.site) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = sim.site, y = prop, fill = sex)) +
  geom_col() +
  theme_linedraw(base_size = 11) +
  ggtitle("Distributions of male and female subjects across simulated sites")
```

Combine plots

```{r}
fig_2 <- ggarrange(
  wmv_centile_fan,
  ggarrange(global_cent_errs_plot, p_diff, ncol=1, nrow=2, heights = c(1.2, 2), common.legend = FALSE, labels=c(NA, "C"), vjust = c(-0.25)),
  ncol=2, nrow=1, 
  labels=c("A", "B"),
  widths=c(1.1, 1)
  #hjust=0.01,
  # align = "hv"
  )

fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Choice of ComBat configuration can induce errors in brain feature centile scores",
               color = "black", hjust = 0, x = 0))
fig_2_annotate
ggsave("figs/figure_2.jpeg", fig_2_annotate, width=11, height=7, units="in")
```


## FIGURE 3

```{r}
blank <- ggplot() + theme_void()

fig_3 <- ggarrange(
  ggarrange(blank, 
            annotate_figure(
              ggarrange(sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.25),
  hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "right"
  )),
  blank,
  ncol=3, nrow=1,
  widths=c(.5, 1, .5)),
  big_violin_plt,
  ncol=1, nrow=2, 
  #labels=c("A", "B"),
  #hjust=0.01,
  # align = "hv",
  heights=c(1, 1.8),
  common.legend = FALSE,
  labels=c("A", "B")
  )
# fig_2

fig_3_annotate <- annotate_figure(fig_3, top = text_grob("Figure 3. Sex-biases in centile score errors across ComBat configurations", color = "black", hjust = 0, x = 0))
fig_3_annotate
ggsave("figs/figure_3.jpeg", fig_3_annotate, width=11.5, height=10, units="in")
```
