---
title: "UKB Simulation Analyses"
output: html_document
date: '2023-10-06'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer, ggsegExtra, ggseg3d, mgcv)

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
source("./cubic_scripts/R_scripts/ranked_welch_tests.R")
# 
# # Enable this universe
# options(repos = c(
#   ggseg = 'https://ggseg.r-universe.dev',
#   CRAN = 'https://cloud.r-project.org'))
# 
# # Install some packages
# install.packages('ggsegExtra')
```

## Load

```{r}
ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")

vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")

aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)

```{r}
drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset)) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb-CN-simsites"))
```

```{r}
#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sex.df <- left_join(sex.df, dk.parc, by="label")
sum.df <- left_join(sum.df, dk.parc, by="label")
```

Simple bar plot of sites
```{r}
ggplot(ukb_df_age.filt) +
  geom_bar(aes(x=sim.site, fill=sex), color="black", alpha=0.6) +
  theme_linedraw(base_size=12)
```

## By Moment

Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods

```{r}
data.csvs <- list.files(path = "data/ukb_to_model", pattern = ".csv", full.names = TRUE)
combined_data <- data.frame()

for (file in data.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file)))
  
  # Bind the data to the combined dataframe
  combined_data <- bind_rows(combined_data, data, .id = "File_ID")
}

table(combined_data$Source_File)
```

```{r}
var.df <- combined_data %>%
  group_by(Source_File, sim.site) %>%
  summarize(across(all_of(pheno_list), var))

var.long <- pivot_longer(var.df, cols=c(-Source_File, -sim.site), names_to = "feature", values_to = "variance")
```

Plot distribution of features' variances across each site

```{r}
var.long %>%
  dplyr::filter(!(feature %in% vol_list_global)) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.4, position="identity") +
  facet_wrap(~sim.site)

var.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" |Source_File == "ukb_CN_data_simsites.csv" |  Source_File == "cf.gamlss_data.csv" )) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.5, position="identity") +
  facet_wrap(~sim.site)
```
How does male vs. female variance across features change by combat type

```{r}
var.by.sex <- combined_data %>%
  group_by(Source_File, sex) %>%
  summarize(across(all_of(pheno_list), var))

var.by.sex.long <- pivot_longer(var.by.sex, cols=c(-Source_File, -sex), names_to = "feature", values_to = "variance")
```

```{r}
var.by.sex.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=sex), alpha = 0.5, position="identity") +
  facet_wrap(~Source_File)
```

Simple t test for sex effects on brain features to see whether # of significantly sex-biased features change across combat configurations

```{r}
sex_ttest_df <- data.frame("Source_File" = levels(combined_data$Source_File))

for(pheno in pheno_list){
  # New column name
  new_col_name <- paste0(pheno, ".p")
  
  #run t-tests
  out <- lapply(split(combined_data, factor(combined_data$Source_File)), function(x)t.test(x[[pheno]] ~ x[["sex"]], paired=FALSE)$p.value)
  stopifnot(names(out) == levels(combined_data$Source_File)) #make sure results are in correct order
  
  #append to dataframe
  sex_ttest_df[[new_col_name]] <- unlist(unname(out))
}
```

```{r}
sex_ttest_long <- sex_ttest_df %>% pivot_longer(!Source_File, values_to="p.value", names_to="pheno") %>%
  mutate(pheno = sub(".p$", "", pheno),
         sig_bfcor = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
                            TRUE ~ NA))
sex_ttest_long %>%
  group_by(Source_File) %>%
  summarize(n_sig_regions = sum(sig_bfcor),
            n_regions = n())
```

Basically no difference across combat types

#### Mu

Does the number of regions w significant age effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

Does the number of regions w significant sex effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())

sum.df %>%
  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.sum_bf.corr),
            n_regions = n())
```

Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "mu" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on mu")
```

Trying with ultra-strict threshold to see if results change:

```{r}
drop1.df <- drop1.df %>%
  mutate(sig.sum.strict = case_when(drop1.pval < (0.05/length(pheno_list)) ~ TRUE,
                            drop1.pval >= (0.05/length(pheno_list)) ~ FALSE,
                            TRUE ~ NA))

drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.sum.strict),
            n_regions = n())
```

No difference

#### Sigma

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset)) +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset), position = "dodge") +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```
Check number of regions w significant sex effects in sigma

```{r}
sex.sigma.counts <- sex.df %>%
  dplyr::filter(term=="sexMale" & parameter == "sigma") %>%
  group_by(dataset, pheno_cat) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
sex.sigma.counts
```
 Plot these results
 
```{r}
ggplot(sex.sigma.counts) +
  geom_col(aes(x=dataset, y=n_sig_regions, fill=dataset)) +
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("# regions with significant effects of male sex on sigma")
```

## Variance
Estimating by 3 different methods:
1. Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).
1.5. getting mean variance exp(mean sigma) (not actually plotted)
2. actual estimated variance at mean(sample age) (andrew's rec)

#### Method 1: exp(sigma)

```{r}
sigma.df <- sum.df %>%
  dplyr::filter(parameter == "sigma") %>%
  mutate(var_est = exp(estimate),
         var_mean = exp(moment.mean),
         beta_ratio = var_est/var_mean)

skim(sigma.df$var_est)

sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_histogram(aes(x=var_est, fill = dataset), color="black", alpha=0.5, position = "identity") +
  theme_linedraw() +
  facet_wrap(~pheno_cat, scales= "free")+
  ggtitle("estimated variance in regions with significant effects of male sex on sigma")
```

What proportion of estimated variance is attributable to sex?

```{r, fig.height=10, fig.width=8}
proportion_beta_plot <- sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = beta_ratio)) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_bw(base_size=15) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork") +
  ggtitle("Proportion of effect of male sex on variance")
```

exp(SexMale beta weight on sigma) -> estimate of male effect on variance

```{r, fig.height=15, fig.width=12}
beta.plot <- sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = var_est)) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
  ggtitle("Beta weight for effect of male sex on variance")

ggsave("figs/male_variance_beta.jpeg", beta.plot, width=12, height=15)
```

```{r, fig.height=5, fig.widt=5}
sigma.df %>%
  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
  mutate(var_est.sig = case_when(sig.sum_bf.corr == TRUE ~ var_est,
                                 sig.sum_bf.corr == FALSE ~ 0)) %>%
  #group_by(dataset) %>%
  ggplot() +
  geom_col(aes(x=dataset, y = var_est.sig, fill=dataset), position="dodge") +
  #scale_fill_gradient2() +
  facet_grid(~pheno) + 
  theme_bw(base_size=12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="none") +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  ggtitle("Beta weight for effect of male sex on global volumes' variance")

sigma.df.glob <- sigma.df %>%
  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale")
table(sigma.df.glob$dataset, sigma.df.glob$sig.sum_bf.corr)
```

#### Method 2: Var at Mean Age
fix age, get diff between male and female pred. variance
```{r}
pred.csvs <- list.files(path = "ukb_basic", pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File))
  
  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

table(combined_var$dataset) #confirm all data's there
#skim(combined_preds)
```

Calculate standardized effects of male sex w/in region as (M var - F var)/F var

```{r}
combined_var <- combined_var %>%
  mutate(sex_std = sex_effect/f.var) #standardized effect of male sex on variance

#compare methods
summary(sigma.df$var_est) #method 1 var
summary(combined_var$sex_effect) #method 2 var

summary(sigma.df$beta_ratio) #method 1 male effect on var
summary(combined_var$sex_std) #method 2 male effect on var
```

```{r}
#join with sigma.df for significance testing
plot.info <- sigma.df %>%
  dplyr::filter(term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
  unique() %>%
  mutate(dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset))
combined_var <- inner_join(combined_var, plot.info)
table(combined_var$dataset) #check no data's lost
```

plot

```{r, fig.height=5, fig.widt=5}
combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_std)) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_bw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, -0.05, 0, 0.05, 0.1)) +
  ggtitle("Standardized effect of male sex on variance")
```

## Centiles

Do centile scores change based on combat version?
```{r}
pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)
  
  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))
  
  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
}

table(combined_preds$dataset)
#skim(combined_preds)
```

#### Magnitude error

Calculate how far centiles are from "truth" (raw data's centile vals), take absolute value for magnitude error

```{r}
cent_diffs <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup()
```

Calculate within feature mean and max abs. centile errors across all subjects

```{r}
mean.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup()

max.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup()
#names(max.cent.diffs)
```

Plot w/in feature average magnitude of centile errors
```{r}
mean.diffs.long <- mean.cent.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
mean.diffs.long <- left_join(mean.diffs.long, dk.parc, by="label")
```

```{r}
mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
ggplot() +
  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
  facet_grid(dataset~pheno_cat)

mean.diffs.long %>%
  dplyr::filter(dataset== "cf.gam" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
  ggplot() +
  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
  scale_x_continuous(labels = label_comma())

mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visualization purposes
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", binwidth=0.1, alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in centile scores across ComBat \nprocessing pipelines for each phenotype")

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
  paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
ggsave("mean_centile_diffs.jpeg", cent.diffs.plot)
```

```{r, fig.height=15, fig.width=12}
mean.diffs.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = centile_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Mean difference in centiles across ComBat configurations")
```

Look at max difference per phenotype

```{r, fig.height=6, fig.width=7}
max.diffs.long <- max.cent.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
  paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
```

```{r}
max.diffs.long %>%
  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity", binwidth = 0.5) +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma())+
  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
```

Test for differences in distribution of subjects' abs(cent error)

```{r}
#get abs val of centile errors for each subject
cent.diffs2 <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "_refA")) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop ref

# cf.gmvdiff <- cent.diffs2 %>%
#   dplyr::filter(dataset == "cf") %>%
#   dplyr::select(diff_GMV)
# cf.lm.gmvdiff <- cent.diffs2 %>%
#   dplyr::filter(dataset == "cf.lm") %>%
#   dplyr::select(diff_GMV) 
# 
# t.test(cf.gmvdiff, cf.lm.gmvdiff, paired=TRUE)
```

plot to see if assumptions for t-tests hold ...
```{r}
ggplot(cent.diffs2) +
  geom_histogram(aes(x=diff_GMV, fill=dataset), alpha = 0.6) +
  theme_bw()
```

looks like.... no. Using welch's t test on centile erro ranks per discussion with Andrew (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`. also updating Anova to nonparametric kruskal-wallace test. **confirm with andrew**

NOTE: "missing" paired comparisons between cf and cf.lm diffs are because diff values are identical between the two methods (i.e. there is literally no difference in accuracy of centile scores between the two)

```{r, warning=FALSE, message=FALSE}
#within each phenotype, iterate paired wilcox rank test between each combo of combat dfs 

#initalize empty dfs to store outputs
sum.df <- data.frame(pheno = character(),
                     term = character(),
                     df = double(),
                     sumsq = double(),
                     meansq = double(),
                     statistic = double(),
                     p.value = double())
t.df <- data.frame("pheno" = character(),
                   "group1" = character(),
                   "group2" = character(),
                   "p.value" = double())
pheno_diff.list <- paste0("diff_", pheno_list)

#run tests
attach(cent.diffs2)
for (pheno in pheno_diff.list) {
  
  #group-wise
  df.aov <- tidy(kruskal.test(cent.diffs2[[pheno]] ~ cent.diffs2[["dataset"]]))
  df.aov$pheno <- pheno
  sum.df <- rbind(sum.df, df.aov)

  #'post-hoc' pairwise t tests
  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=cent.diffs2[[pheno]], g=cent.diffs2[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  t.df <- rbind(t.df, df.pairwise.t)
}
detach()
```

```{r}
#apply fdr correction
sum.df <- sum.df %>%
  dplyr::filter(term != "Residuals") %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

summary(sum.df$p.val_fdr)


t.df <- t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("diff_", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))
table(t.df$comp)

t.df %>%
  group_by(comp) %>%
   summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n(),
            min_pval = min(p.val_fdr),
            max_pval = max(p.val_fdr))
```

#### Sex-Biases

Do the different combat methods introduce any kind of bias in centile scores?

```{r}
mean.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

mean.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")


max.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_diff = max_centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")
```

No shift in abs. differences, looking at overall diff.

```{r}
#get centile errors for each subject - not abs val
cent.diffs3 <- combined_preds %>%
  dplyr::filter(!str_detect(dataset, "_refA")) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
  ungroup() %>%
  dplyr::filter(dataset != "ukb_CN") #drop ref
saveRDS(cent.diffs3, file="ukb_basic/cent_diffs.RDS")
```

Look at w/in subject maximum error across all feature's centiles & compare across harmonization methods

```{r}
max.cent.diffs.by.sex2 <- cent.diffs3 %>%
  group_by(dataset, sex) %>%
  summarize_at(vars(starts_with("diff")), list(max_value = ~.[which.max(abs(.))])) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_diff = max_centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales="free")
```

all methods are fairly consistently underestimating male centiles and overestimating female centiles.

```{r}
mean.cent.diffs.by.sex2 <- cent.diffs3 %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

mean.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales="free")
```

test of mean diff. between sex w/in each dataset - again need to update w nonparametric method **ask andrew about this**
```{r}
t.test.by.sex <- mean.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  group_by(dataset) %>%
  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
            p_value = t.test(centile_diff ~ sex)$p.value) %>%
  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
                             p_val_fdr.corr >= 0.05 ~ FALSE))

t.test.by.sex
```

Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.

t tests of sex-biases in centiles w/in each feature

```{r, warning=FALSE, message=FALSE}
# within each df, within each phenotype, t-test between male and female diffs

#initalize empty dfs to store outputs
t.df.sex <- data.frame("pheno" = character(),
                   "dataset" = character(),
                   "p_value" = double()
                   )

for (pheno in pheno_diff.list){
    df.sex.test <- cent.diffs3 %>%
        group_by(dataset) %>%
        summarize(p_value = rank.welch.t.test(.data[[pheno]][sex == "Male"], .data[[pheno]][sex == "Female"])$p.value) %>%
        ungroup()
    df.sex.test$pheno <- pheno
    t.df.sex <- rbind(t.df.sex, df.sex.test)
}
```

```{r}
#apply fdr correction
t.df.sex <- t.df.sex %>%
  mutate(p.val_fdr = p.adjust(p_value, method="fdr", n = length(p_value)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("diff_", "", pheno))

t.df.sex %>%
  group_by(dataset) %>%
   summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n())
```

### Extremeness Bias

Does a person's centile error depend on their true centile value (e.g., very low or very high centiles more or less likely to have errors)

do some plotting

```{r}
ggplot(cent.diffs3, aes(x=GMV, y=diff_GMV, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  facet_wrap(~dataset)

ggplot(cent.diffs3, aes(x=WMV, y=diff_WMV, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  facet_wrap(~dataset, scales="free")
```

Testing on MAGNITUDE of error, not the errors themselves!

```{r}
ggplot(cent.diffs2, aes(x=GMV, y=diff_GMV, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  facet_wrap(~dataset)

ggplot(cent.diffs2, aes(x=WMV, y=diff_WMV, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  facet_wrap(~dataset, scales="free")
```

*ask andrew about how best to test - would GAM be too flexible? use polynomials?*

## Z-Score
calc how much z-scores change

```{r}
z_diffs <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup()
```

```{r}
#mean
mean.z.diffs <- z_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup()
mean.z.long <- mean.z.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno = sub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
mean.z.long <- left_join(mean.z.long, dk.parc, by="label")

#max
max.z.diffs <- z_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup()
max.z.long <- max.z.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno = sub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))
```

```{r}
mean.z.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=z_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
```

```{r, fig.height=15, fig.width=12}
mean.z.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = z_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Mean difference in z scores across ComBat configurations")
```

```{r}
max.z.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=max_z_abs.diff, fill=dataset), color="black", binwidth = 0.025, alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
```

```{r, fig.height=15, fig.width=12}
max.z.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = max_z_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Maximum difference in z scores across ComBat configurations")
```


Test for differences in distribution

```{r, warning=FALSE, message=FALSE}
#within each phenotype, iterate paired t test between each combo of combat dfs 

z_diffs.filt <- z_diffs %>%
  dplyr::filter(!str_detect(dataset, "_refA")) %>%
  dplyr::filter(dataset != "ukb_CN")

#initalize empty dfs to store outputs
z.sum.df <- data.frame(pheno = character(),
                     term = character(),
                     df = double(),
                     sumsq = double(),
                     meansq = double(),
                     statistic = double(),
                     p.value = double())
z.t.df <- data.frame("pheno" = character(),
                   "group1" = character(),
                   "group2" = character(),
                   "p.value" = double())

z.pheno_abs.diff.list <- paste0("abs.diff_", pheno_list, ".z")

#run tests
attach(z_diffs.filt)
for (pheno in z.pheno_abs.diff.list) {
  df.aov <- tidy(kruskal.test(z_diffs.filt[[pheno]] ~ z_diffs.filt[["dataset"]]))
  df.aov$pheno <- pheno
  z.sum.df <- rbind(z.sum.df, df.aov)

  df.pairwise.t <- tidy(pairwise.rank.welch.t.test(x=z_diffs.filt[[pheno]], g=z_diffs.filt[["dataset"]], paired = TRUE, p.adj = "none"))
  df.pairwise.t$pheno <- pheno
  z.t.df <- rbind(z.t.df, df.pairwise.t)
}
detach()
```

```{r}
#apply fdr correction
z.sum.df <- z.sum.df %>%
  dplyr::filter(term != "Residuals") %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno_list)))

summary(z.sum.df$p.val_fdr)

z.t.df <- z.t.df %>%
  mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(pheno)),
         sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                             p.val_fdr >= 0.05 ~ FALSE),
         pheno = sub("abs.diff_", "", pheno),
         pheno = sub(".z", "", pheno)) %>%
  unite(comp, c("group1", "group2")) %>%
  mutate(comp = as.factor(comp))

z.t.df %>%
  group_by(comp) %>%
   summarize(n_sig_regions = sum(sig_fdr),
            n_compare = n(),
            min.pval = min(p.val_fdr),
            max.pval = max(p.val_fdr))
```

##### Sex-Biases

Test for sex-specific differences in mean z-score offset

```{r}
z_diffs2 <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(ends_with(".z"), ~ . - first(.), .names = "diff_{.col}")) %>%
  ungroup()
#write out
saveRDS(z_diffs2, file = "ukb_basic/z_diffs.RDS")

mean.z.diffs.by.sex2 <- z_diffs2 %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="z_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno = gsub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))
```
t-test

```{r}
#check for skew
ggplot(z_diffs2) +
  geom_histogram(aes(x=diff_sGMV.z, fill=dataset))
```


```{r}
t.test.by.sex_z <- mean.z.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  group_by(dataset) %>%
  summarise(t_statistic = t.test(z_diff ~ sex)$statistic,
            p_value = t.test(z_diff ~ sex)$p.value) %>%
  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
                             p_val_fdr.corr >= 0.05 ~ FALSE))

t.test.by.sex_z
```


Same results as centiles.

### Extremeness Bias

Plot error

```{r}
ggplot(z_diffs2, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)

ggplot(z_diffs2, aes(x=WMV.z, y=diff_WMV.z, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)
```

plot magnitude error (abs val)

```{r}
ggplot(z_diffs, aes(x=GMV.z, y=abs.diff_GMV.z, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  facet_wrap(~dataset)

ggplot(z_diffs, aes(x=WMV.z, y=abs.diff_WMV.z, color=sex)) +
  geom_point(alpha=0.1) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)
```

Try using gam to test associations (start with just GMV)

```{r}
m1 <- gam(abs.diff_GMV.z ~ s(GMV.z, by=dataset) + dataset, data=z_diffs)
summary(m1)

#test for interactive sex effects - split df by dataset and smooth by sex
for (dset in c("cf", "cf.gam", "cf.gamlss", "cf.lm")) {
  print(dset)
  df <- z_diffs %>%
    dplyr::filter(dataset == dset)
  
  m2 <- gam(abs.diff_GMV.z ~ s(GMV.z, by=as.factor(sex)) + as.factor(sex) + s(GMV.z), data=df)
  print(summary(m2))
}
```

Pivot data long and run gam to test assoc. across all features
```{r}
dataset_test_list <- c("cf", "cf.gam", "cf.gamlss", "cf.lm")
pheno_list.z <- paste0(pheno_list, ".z")

z_diffs.long <- z_diffs %>%
  dplyr::filter(dataset %in% dataset_test_list) %>%
  mutate(dataset = droplevels(dataset)) %>%
  dplyr::select(c(participant, sex, age_days, dataset, Source_File), starts_with("abs.diff_")) %>%
  pivot_longer(starts_with("abs.diff_"), names_to="pheno", values_to="z_abs.diff", names_prefix = "abs.diff_") 

z_long <- z_diffs %>%
  dplyr::filter(dataset %in% dataset_test_list) %>%
  mutate(dataset = droplevels(dataset)) %>%
  dplyr::select(c(participant, sex, age_days, dataset, Source_File, pheno_list.z)) %>%
  pivot_longer(cols=all_of(pheno_list.z), names_to="pheno", values_to="z_score")

z_long.full <- inner_join(z_diffs.long, z_long)
```

```{r}
gam.ext_m1 <- gam(z_abs.diff ~ s(z_score, by=dataset) + dataset, data=z_long.full)
summary(gam.ext_m1)
```

```{r}
#test for interactive sex effects - split df by dataset and smooth by sex 
for (dset in c("cf", "cf.gam", "cf.gamlss", "cf.lm")) {
  print(dset)
  df <- z_long.full %>%
    dplyr::filter(dataset == dset)
  
  m2 <- gam(z_abs.diff ~ s(z_score, by=as.factor(sex)) + as.factor(sex) + s(z_score), data=df)
  print(summary(m2))
}
```
