---
title: "UKB Models"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
date: '2023-09-23'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

## Load Data

Load UKB Data from Jakob

```{r}
cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
#table(cn_data$study)
names(cn_data)
```

## Data Org

Filter

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         TBV=(sGMV+GMV+WMV+Ventricles),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale * age_days) %>%
  na.omit() #1 subj has 1 missing val

skim(ukb_df)
n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
```

Write lists of phenotypes

```{r}
col_list <- colnames(ukb_df)
pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
length(pheno_list) #total number of models to run
saveRDS(pheno_list, file="R_scripts/pheno_list.rds")

#split by measure type (Vol, SA, CT)
vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
saveRDS(vol_list_global, file="R_scripts/vol_list_global.rds")

vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
saveRDS(vol_list_regions, file="R_scripts/Vol_list_regions.rds")

sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
writeLines(sa_list, con = "pheno_lists/SA_list.txt")
saveRDS(sa_list, file="R_scripts/SA_list.rds")

ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
writeLines(ct_list, con = "pheno_lists/CT_list.txt")
saveRDS(ct_list, file="R_scripts/CT_list.rds")
```

Add brainwide phenotypes!
Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT

```{r}
vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]

ukb_df_totals <- ukb_df %>%
  mutate(Vol_total=rowSums(dplyr::select(., .dots=all_of(c(vol_list_regions)))),
         SA_total=rowSums(dplyr::select(., .dots=all_of(c(sa_list)))),
         CT_total=rowSums(dplyr::select(., .dots=all_of(c(ct_list)))))

#write data
write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
```


```{r}
ggplot(data=ukb_df, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
ggplot(data=ukb_df, aes(x=log_age)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
  geom_point(aes(color=sex), alpha=0.6)
ggplot(data=ukb_df, aes(x=age_days, y=GMV)) +
  geom_point(aes(color=sex), alpha=0.6)
```

### Condense age range

Random male subjects are younger than females - going to drop because having issues estimating total brain metrics in young women where there's no data

```{r}
ukb_df_totals %>%
  dplyr::filter(age_yrs <= 49) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)

ukb_df_totals %>%
  dplyr::filter(age_yrs >= 79) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
```

Restricting to ages 50-80, since that's when there's roughly ~100 ppts per sex

```{r}
ukb_df_age.filt <- ukb_df_totals %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
fwrite(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
```

## EDA

```{r}
#percent female
r <- table(ukb_df_age.filt$sex)

(r["Female"]/n_unique(ukb_df_age.filt$participant))*100
summary(ukb_df_age.filt$age_yrs)
(table(ukb_df_age.filt$fs_version)["FS6_T1T2"])/(n_unique(ukb_df_age.filt$participant))*100
```

Going to stick with age_days since distribution is fairly normal, not lifespan data

```{r}
ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV

ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")
```

Look at variances

center
```{r}
ukb_centered <- ukb_df_totals %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

```{r}
# Generate histograms (mean-centered w/in sex)
global_plots <- plot.hist.by.sex(ukb_centered, vol_list_global)

# Print or display the histograms
for (pheno in vol_list_global) {
  print(global_plots[[pheno]])
}

ukb_centered %>%
  dplyr::filter(sex == "Male") %>%
  dplyr::select(GMV) %>%
  var()

ukb_centered %>%
  dplyr::filter(sex == "Female") %>%
  dplyr::select(GMV) %>%
  var()

ukb_centered %>%
  #dplyr::filter(sex == "Male") %>%
  dplyr::select(GMV) %>%
  var()
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_vol_plots <- plot.hist.by.sex(ukb_centered, vol_list_regions)

# Print or display the histograms
for (pheno in vol_list_regions) {
  print(region_vol_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_sa_plots <- plot.hist.by.sex(ukb_centered, sa_list)

# Print or display the histograms
for (pheno in sa_list) {
  print(region_sa_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_ct_plots <- plot.hist.by.sex(ukb_centered, ct_list)

# Print or display the histograms
for (pheno in ct_list) {
  print(region_ct_plots[[pheno]])
}
```

## Simulate Sites
Dividing each subject into 3 sites (A, B, and C) with M:F ratios of 1:1, 1:4, and 4:1

```{r}
set.seed(1010101)
sim.site.list <- c("Site_A", "Site_B", "Site_C")
n_female <- table(ukb_df_age.filt$sex)["Female"]
n_male <- table(ukb_df_age.filt$sex)["Male"]

# Sample probabilities
male_prob <- c(0.33, 0.0825, 0.5875)
female_prob <- c(0.33, 0.5875, 0.0825)

# Separate samples for males and females
sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)

ukb_df_sites <- ukb_df_age.filt %>%
  mutate(sim.site = as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))

#write out
write_csv(ukb_df_sites, file="data/ukb_to_model/ukb_CN_data_simsites.csv")
```

```{r}
table(ukb_df_sites$sim.site)
table(ukb_df_sites$sim.site, ukb_df_sites$sex)
sum(is.na(ukb_df_sites$sim.site)) #all subj assigned to sites
anova(lm(age_days ~ sim.site, ukb_df_sites))
```

Need to come up with cleaner way to do this...

```{r}
site_sum_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  summarise(proportion_male = mean(sex == "Male"),
            min_age = min(age_days), max_age = max(age_days), mean_age = mean(age_days), 
            var_gmv = var(GMV), var_wmv=var(WMV), var_sgmv=var(sGMV), var_ven = var(Ventricles),
            sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))
```

Center and visualize distributions

```{r}
ukb_sites_sex.centered <- ukb_df_sites %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

```{r}
# Generate histograms (mean-centered w/in sex)
global_site_plots <- plot.hist.by.sex(ukb_sites_sex.centered, vol_list_global, facet_fac = "sim.site")

# Print or display the histograms
for (pheno in vol_list_global) {
  print(global_site_plots[[pheno]])
}
```

Look at uncentered distributions across sites
```{r}
ggplot(ukb_df_sites, aes(x=GMV)) + 
 geom_histogram(aes(y=after_stat(density)))+
 geom_density(alpha=.2) +
  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV))))
  

ggplot(ukb_df_sites, aes(x=GMV)) + 
 geom_histogram(aes(y=after_stat(density)))+
 geom_density(alpha=.2) +
  geom_vline(aes(xintercept=(mean(GMV)+sd(GMV)))) +
  geom_vline(aes(xintercept=(mean(GMV)-sd(GMV)))) +
  facet_wrap(~sim.site)
```

```{r}
centered.site_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()

centered.site_sum_df <- centered.site_df %>%
  group_by(sim.site) %>%
  summarise(sd_below.gmv = mean(GMV) - sd(GMV), sd_above.gmv = mean(GMV) + sd(GMV))

ggplot() + 
 #geom_histogram(aes(x=centered.site_df$GMV, y=after_stat(density), fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha = .2)+
 geom_density(aes(x=centered.site_df$GMV, fill=centered.site_df$sim.site, color=centered.site_df$sim.site), alpha=.4) +
  geom_vline(aes(xintercept=centered.site_sum_df$sd_below.gmv, color=centered.site_sum_df$sim.site)) +
  geom_vline(aes(xintercept=centered.site_sum_df$sd_above.gmv, color=centered.site_sum_df$sim.site))
```

## Simulate Sites - Permutations

```{r}
set.seed(1010101)
sim.site.list <- c("Site_A", "Site_B", "Site_C")
n_female <- table(ukb_df_age.filt$sex)["Female"]
n_male <- table(ukb_df_age.filt$sex)["Male"]

# Sample probabilities
male_prob <- c(0.33, 0.0825, 0.5875)
female_prob <- c(0.33, 0.5875, 0.0825)

# Create a list to store data frames
df_list <- list()

for(i in 1:5) {
  new_col_name <- paste0("SimSite", i)
  # Separate samples for males and females
  sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
  sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)

  new_df <- ukb_df_age.filt %>%
    mutate(!!new_col_name := as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))%>%
    dplyr::select(!!new_col_name)
  
  # Store the data frame in the list
  df_list[[i]] <- new_df
}

# Combine data frames into a single data frame
ukb_df_sites_perm <- bind_cols(ukb_df_age.filt, !!!df_list)

# View the final dataframe
#names(ukb_df_sites_perm)
```

```{r}
sim.perm.long <- ukb_df_sites_perm %>%
  dplyr::select(sex, age_days, contains("SimSite")) %>%
  pivot_longer(!c(sex, age_days), names_to = "permutation", names_prefix = "SimSite", values_to = "site")

prop <- sim.perm.long %>%
  group_by(permutation, site, sex) %>%
  summarize(Count = n(), avg_age = mean(age_days)) %>%
  group_by(permutation, site) %>%
  mutate(Proportion = Count / sum(Count))

summary.prop.male <- prop %>%
  dplyr::filter(sex=="Male") %>%
  group_by(site) %>%
  summarize(avg_prop = mean(Proportion), min_prop = min(Proportion), max_prop = max(Proportion))

summary.prop.male <- prop %>%
  group_by(site, sex) %>%
  summarize(avg_age_days = mean(avg_age), min_age = min(avg_age), max_age = max(avg_age))
```

```{r}
#write out
write_csv(ukb_df_sites_perm, file="data/ukb_CN_data_simsites_perm.csv")
```
