---
title: "lifespan_eda"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed, ggrain, ggpubr)

base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))


source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source(paste0(base_path, "/cubic_scripts/R_scripts/gamlss_helper_funs.R"))
```

Load data from Jakob - updated with euler numbers and corrected VETSA study
```{r}
cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE, na.strings="")
```

Plot full samples for supplemental figures looking at imbalanced demographics across sites

```{r}
#just get first scan
cn.df_first <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice_head(n=1) %>%
  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&"))),
         age_yrs = (age_days-280)/365.25) %>%
  ungroup() %>%
  group_by(site_full) %>%
  filter(n() > 1) #keep sites with at least 2 ppl
```

check for sex differences in sites

```{r}
sfig_full_sample <- ggplot(cn.df_first) +
  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site, study, sep="&"), fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(y = ggh4x::guide_axis_nested(delim="&")) +
  theme_linedraw() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    ggh4x.axis.nesttext.y = element_text(size=6),
    ggh4x.axis.nestline = element_line(linetype = 1),
    legend.position="right"
  ) +
  geom_vline(xintercept=.5, color="lightgray", linetype="dashed") +
  labs(y="Data Collection Site", x="Sample Composition", fill="Sex")
ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=6, height=14, units="in")
```

look into a couple studies specifically

```{r}
plot_by_sex <- function(df) {
  df_prepared <- df %>%
    group_by(site, sex) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(site) %>%
    mutate(percentage = count / sum(count)) %>%
    ungroup()
  
  ggplot(df_prepared, aes(x = site, y = percentage, fill = sex)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(
      aes(label = count, group = sex),
      position = position_stack(vjust = 0.5),
      color = "white",
      size=2.5
    ) +
    scale_y_continuous(labels = scales::percent_format()) +
    theme_linedraw() +
    xlab("Site") +
    ylab("Percentage") +
    labs(fill="Sex") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

#good balance
p1<- cn.df_first %>% 
  filter(study == "ABCD") %>% 
  plot_by_sex() + labs(title="ABCD")
  

p2 <- cn.df_first %>% 
  filter(study == "UKB") %>% 
  plot_by_sex() + labs(title="UK Biobank")

#bad balance
p3 <- cn.df_first %>% 
  filter(study == "cVEDA") %>% 
  plot_by_sex() + labs(title="cVEDA")

p4 <- cn.df_first %>% 
  filter(study == "ADNI") %>% 
  plot_by_sex() + labs(title="ADNI")

p5 <- cn.df_first %>% 
  filter(study == "abide1") %>% 
  plot_by_sex() + labs(title="ABIDE 1")

p6 <- cn.df_first %>% 
  filter(study == "ADHD200") %>% 
  plot_by_sex() + labs(title="ADHD200")

sfig_samples <- ggarrange(p1, p2, p3, p4, p5, p6, 
                          ncol=2, nrow=3, 
                          common.legend = TRUE, legend="right",
                          align="v",
                          labels = "AUTO")
ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, 
       width=15, height=10, units="in")
```

look at age variability

```{r}
sfig_full_sample_age <- ggplot(cn.df_first) +
  geom_boxplot(aes(x=age_yrs, y=interaction(site, study, sep="&"))) +
  guides(y = ggh4x::guide_axis_nested(delim="&")) +
  theme_linedraw() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    ggh4x.axis.nesttext.y = element_text(size=6),
    ggh4x.axis.nestline = element_line(linetype = 1),
    axis.title.y=element_blank()
  ) +
  labs(x="Age (Years Post-Birth)")
ggsave(paste0(base_path, "/figs/sfig_full_sample_age.jpeg"), sfig_full_sample_age, width=6, height=14, units="in")

sitewise_full_plot <- ggarrange(sfig_full_sample, sfig_full_sample_age,
          ncol=2, nrow=1, align = "hv",
          common.legend = TRUE, legend = "bottom",
          labels="AUTO")
ggsave(paste0(base_path, "/figs/sitewise_full_plot.jpeg"), sitewise_full_plot, 
       width=15, height=15, units="in")
```


```{r}
plot_by_age <- function(df) {
  ggplot(df, aes(y = age_yrs, x=site)) +
    geom_boxplot() +
    theme_linedraw() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    xlab("Site") +
    ylab("Age")
}

p1.age <- cn.df_first %>% 
  filter(study == "ABCD") %>% 
  plot_by_age() + labs(title="ABCD")

p2.age <- cn.df_first %>% 
  filter(study == "UKB") %>% 
  plot_by_age() + labs(title="UK Biobank")

#bad balanced
p3.age <- cn.df_first %>% 
  filter(study == "EMBARC") %>% 
  plot_by_age() + labs(title="EMBARC")

p4.age <- cn.df_first %>% 
  filter(study == "ADNI") %>% 
  plot_by_age() + labs(title="ADNI")

p5.age <- cn.df_first %>% 
  filter(study == "abide1") %>% 
  plot_by_age() + labs(title="ABIDE 1")

p6.age <- cn.df_first %>% 
  filter(study == "ADHD200") %>% 
  plot_by_age() + labs(title="ADHD200")

sfig_samples.age <- ggarrange(p1.age, p2.age, p3.age, p4.age, p5.age, p6.age, 
                          ncol=2, nrow=3,
                          align="v",
                          labels = "AUTO")
ggsave(paste0(base_path, "/figs/sfig_subsamples_age.jpeg"), sfig_samples.age, 
       width=15, height=10, units="in")
```

```{r}
#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
nrow(cn.df) - nrow(cn.df.unique)
```

```{r}
skim(cn.df.unique)
levels(cn.df.unique$study)
n_unique(cn.df.unique$study)
```

euler # noted as surfholes

## Visualize Features
plot global brain features for basic qc - check again post-combat

```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=Ventricles, color=sex))
```

log-scaled
```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=Ventricles, color=sex))
```

## Curate

Remove infants (anyone younger than 3 yrs)

```{r}
cn.df.noinf <- cn.df.unique %>%
  dplyr::filter(age_days >= (365.25*3+280)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df.unique) - nrow(cn.df.noinf)
n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

Visualizing euler distributions
```{r}
cn.df.noinf %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes))
```

Removing subjects w/o surfholes, subjects with surfholes values > 2 MADs above study median (same as sensitivity analyses in LBCC)

```{r}
cn.df.eulerfilt <- cn.df.noinf %>%
  dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data

study_mad <- cn.df.eulerfilt %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(surfholes),
            # study_euler_madconst1 = mad(surfholes, constant=1),
            study_euler_median = median(surfholes)) %>%
  mutate(euler_threshold = study_euler_median+(2*study_euler_mad))

left_join(cn.df.eulerfilt, study_mad) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes)) +
  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red")

left_join(cn.df.eulerfilt, study_mad) %>%
  mutate(scan_pass = ifelse(surfholes <= euler_threshold, "yes", "no")) %>%
  group_by(study) %>%
  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
  summarise(max_fail = max(fail_percent),
            min_fail=min(fail_percent),
            mean_fail=mean(fail_percent))
```

using default mad() constant, looks better

```{r}
cn.df.eulerfilt2 <- left_join(cn.df.eulerfilt, study_mad) %>%
  dplyr::filter(surfholes <= euler_threshold)
```

Removed `r nrow(cn.df.noinf) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain


Keep studies w missing site info and just treat as single-site

```{r}
#clean up site NAs
cn.df.imp1 <- cn.df.eulerfilt2 %>%
  mutate(site = gsub(" ", "_", trimws(as.character(site))),
         study = gsub(" ", "_", trimws(as.character(study)))) %>%
  mutate(site = ifelse(is.na(site), study, paste(study, site, sep="_"))) 
skim(cn.df.imp1)
```

```{r}
#drop NAs listwise
cn.df.imp <- cn.df.imp1%>%
  dplyr::select(!c(INDEX.ID)) %>%#drop index ID for filtering
  na.omit()

skim(cn.df.imp)

nrow(cn.df.imp1) - nrow(cn.df.imp)
n_unique(cn.df.imp$site)
n_unique(cn.df.imp$study)
# table(cn.df.imp$site, cn.df.imp$study)
# sum(is.na(cn.df.imp$site))
```

##Demographics

```{r}
#sample size, sex
length(unique(cn.df.imp$participant))
nrow(cn.df.imp) == length(unique(cn.df.imp$participant))

table(cn.df.imp$sex)

table(cn.df.imp$sex)["Female"]/nrow(cn.df.imp)

#studies, sites
length(unique(cn.df.imp$study))
length(unique(cn.df.imp$site))

#age
min(cn.df.imp$age_days)/365.25-(280/365.25)
max(cn.df.imp$age_days)/365.25-(280/365.25)
mean(cn.df.imp$age_days)/365.25-(280/365.25)

abcd <- cn.df.imp %>%
  dplyr::filter(study=="ABCD")
length(unique(abcd$site))
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(study, age_days), y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

Write out

```{r}
fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))

#save names with site (since not saved correctly from centile outputs)
study_site_key <- cn.df.imp %>%
  dplyr::select(study, site, country, fs_version) %>%
  unique()
n_unique(study_site_key$study) == n_unique(cn.df.imp$study)
n_unique(study_site_key$site) == n_unique(cn.df.imp$site)
fwrite(study_site_key, file=paste0(base_path, "/data/lifespan_study_site_key.csv"))
```

###Plot

Clean up names, order by age
```{r}
#reorder studies by age
age.df <- cn.df.imp %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_days)) %>%
  arrange(mean_age)
study_names_clean <- gsub("_", " ", unique(age.df$study))

final_df <- cn.df.imp %>% 
  mutate(age_years=(age_days-280)/365.25,
         study = gsub("_", " ", study)) %>%
  mutate(study = factor(study, levels=study_names_clean, ordered=TRUE))
```

Sample plot
```{r}
sample_boxplot <- final_df %>%
  #mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=study, x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
  labs(color="Sex", y = "Study")

#overall sample
full_boxplot <- final_df %>%
  mutate(ylab = "   Full Sample") %>%
  ggplot(aes(y=ylab, x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x = "Age (Years)", y="")

sample_barplot <- final_df %>%
  #mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(fill="Sex")

full_barplot <- final_df %>%
  mutate(ylab = "Full Sample") %>%
  ggplot() +
  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
  labs(fill="Sex", x = "Sex")

ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          align = "hv", 
          common.legend = TRUE,
          widths=c(2, 1),
          heights=c(10, 1))
```

look at country info

```{r}
sample_boxplot2 <- final_df %>%
  ggplot(aes(y=age_years, x=study, fill=country)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.text.x = element_blank(), legend.position = "right") +
  labs(fill="Country", y = "Age (Years)")  +
  paletteer::scale_fill_paletteer_d("ggthemes::Tableau_20")
```

```{r}
blank <- ggplot() + theme_void()

#overall sample
full_boxplot2 <- final_df %>%
  mutate(xlab = "Full Sample ") %>%
  ggplot(aes(x=xlab, y=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
  labs(color="Sex", x=" ")

sample_barplot2 <- final_df %>%
  ggplot() +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
  labs(fill="Sex", x = "Study", y="Percent Sex")

full_barplot2 <- 
  final_df %>%
  mutate(xlab = "  Full Sample  ") %>%
  ggplot() +
  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
  labs(color="Sex", y = "Age (Years)", y="")

full_rain <- final_df %>%
  ggplot(aes(x=sex, y=age_years, fill=sex)) +
  geom_rain(alpha=.4) +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), 
        axis.title.x = element_blank(), axis.text.y = element_text(size=6),
        legend.position = "none") +
  labs(y="Age (Years)")
```

```{r}
country_leg <- as_ggplot(get_legend(sample_boxplot2))
sex_leg <- as_ggplot(get_legend(sample_barplot2))

fig5A <- ggarrange(sample_boxplot2, country_leg,
          sample_barplot2,  sex_leg,
          nrow=2, ncol=2, 
          #align = "v", 
          common.legend = FALSE,
          legend = "none",
          widths=c(11, 1),
          heights=c(1.9, 1)
          )
```

```{r}
fig5_AB <- ggarrange(fig5A, full_rain,
          nrow=1, ncol=2, 
          legend = "none",
          widths=c(3, 1),
          labels = "AUTO")

saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
```